### 2017-11-13 Manuscript/carbonStabilizationMangroves.tex
Added manuscript files

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended, referee]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{here}
\usepackage{upgreek}
\usepackage{wasysym}
\usepackage{natbib}
\usepackage{lineno}
\usepackage{url}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Wetlands Ecology and Management}
%
\begin{document}

\title{Stabilization of carbon in mineral soils from mangroves of the Sin\'{u} river delta, Colombia%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

\titlerunning{Mineral C in mangrove soils}        % if too long for running head

\author{Heidi V\"olkel         \and
        Jhoanata M. Bolivar  \and Carlos A. Sierra
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{H. V\"olkel and C. A. Sierra \at
              Max Planck Institute for Biogeochemistry, Hans-Kn\"oll Str. 10, 07745 Jena, Germany \\
              %Tel.: + 49 3641 ?\\
              %Fax: +123-?\\
              \email{csierra@bgc-jena.mpg.de}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           J. M. Bolivar \at
              Research Center on Ecosystems and Global Change Carbono \& Bosques, Medell\'{i}n, Colombia 
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\linenumbers

\begin{abstract}
Mangrove forests of the Sin\'{u} river delta in Cispat\'{a} bay, Colombia, show large differences in soil carbon storage between fringe (oceanic) and basin (estuarine) mangroves. We were interested in testing whether these differences in soil carbon are associated with sediment transport processes or whether most of the carbon is produced in situ within the mangrove system. Given past sedimentation dynamics of the Sin\'u river, we hypothesized that a large portion of soil carbon in basin mangroves is due to sedimentation. 
We determined total organic carbon content (TOC) of 661 $\pm$ 116 MgC ha$^{-1}$ for basin soils up to a sampling depth of 1 m, and  320 $\pm$ 60 MgC ha$^{-1}$ for fringe soils up to 80 cm depth (maximum soil depth for fringe soils).  Using analyses of mineralogy (Al- and Fe-oxides, clay minerals) as well as isotopic analyses of carbon ($\updelta^{13}$C), the origin of the sediments and their carbon was determined. We found that basin soils in Cispat\'{a} bay show similar mineralogical composition than those of fluvial sediments, but the carbon concentration of river sediments was close to zero.  
Given the large capacity of the Fe and Al oxides in clay minerals to store dissolved carbon, and that the isotopic composition of the carbon is mostly of plant origin, we concluded contrary to our initial hypothesis that the carbon stored in basin mangrove soils are produced in situ. The deposited fluvial sediments do play an important role for carbon storage, but mostly in providing binding surfaces for the stabilization of organic carbon.

%Insert your abstract here. Include keywords, PACS and mathematical
%subject classification numbers as needed.
\keywords{soil organic carbon \and stable isotopes \and iron and aluminum oxides \and soil mineralogy \and estuarine ecosystems}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Although mangroves are ecosystems with some of the largest levels of carbon storage on earth \citep{Donato2011, Alongi2012}, there are  large variations on the amount of C stored in various systems, particularly for soils. For instance, mangrove ecosystems dominated by \emph{Rhizophora} spp. in Peninsular Malaysia store between 479 to 2205 Mg C ha$^{-1}$ in the belowground and soil pool \citep{Alongi2012}, and similar levels of variability have been observed at other sites \citep{Jardine2014}. It is unclear however, what are the main determinants of observed differences in soil C storage across diverse mangrove systems. It is possible that differences in soil carbon storage are due to differences in the level of productivity of different mangrove systems, or due to other external sources such as sediment transport.

This large degree of spatial variability in soil carbon storage is also well expressed in the mangrove forests of Cispat\'{a} bay, Colombia. These mangroves consist of two main forest types: basin and fringe systems, 
which show large differences in terms of soil carbon stocks between them. A previous study \citep{Bolivar2015} showed that for basin mangroves the total organic carbon storage (TOC) is around 740$\pm$40 Mg C ha$^{-1}$, while for the fringe mangroves this value is only 95$\pm$9 Mg C ha$^{-1}$. These numbers, particularly for the basin mangroves, are in the upper range of values observed for other systems \citep{Donato2011, Alongi2012, Jardine2014}.

%An example for this is Gazi bay in Kenya, where the mangrove soil shows an average carbon content of only 36.55 Mg C ha$^{-1}$ at a sampling depth of 60 cm \citep{tamooh2008}. Another example is Okinawa Island in Japan, where the TOC content is 57.30 Mg C ha$^{-1}$ at a sampling depth of 1 m \citep{khan2007}. \par
It is unknown whether the high levels of soil carbon in Cispat\'{a} bay are due to the intrinsically high levels of productivity of these systems or whether this soil carbon has an external source such as transport of fluvial or marine sediments.  Given that mangrove productivity is high and decomposition in water saturated soils is slow, carbon stored in these systems may not have any external origin \citep{lacerda}. However, it is also possible that sediments in the delta region may have been deposited by the Sin\'u river given that before 1938 it discharged in the current mangrove area \citep{suarez2004}, in which case the carbon stored in these sediments may have its origins in soils from the northern Andean mountains.  Alternatively, the carbon in these sediments may have been transported by marine tides over the Caribbean.

Here, our main objective was to determine the origin of the relatively high carbon levels in these soils using elemental and isotopic analyses of carbon as well as analyses of the soils' mineralogy in concert with Al- and Fe-oxide measurements. Analyses of stable isotopes are particularly useful to identify the origin of carbon in soils of coastal areas \citep{Bouillion, Spohn2012, Spohn2013}. In particular, we expect that: 1) the isotopic composition ($\updelta ^{13}$C) of soil carbon provides information on whether riverine sediments are a main source of C in Cisipat\'a bay, Colombia; and  2) the mineralogical composition of the sediments provides additional information on the origin of the soil carbon and the potential that it is mostly stabilized on the surface of Fe- and Al-oxides. Our main hypothesis is that soils in the basin mangroves are composed mainly by sediments transported by the Sin\'u river and deposited in the delta region, therefore explaining the relatively large values of TOC stored in these soils.

%\begin{enumerate}
%    \item Sediments in the mangrove forest of Cispat\'{a} bay show significant differences in carbon stocks between basin- and fringe mangroves
%    \item Additional entered carbon has its origin both in alluvial deposits through tidal flooding in Cispat\'{a} bay and in sedimentation via the Sin\'{u} river delta estuary
%    \item Transformed Al and Fe oxides and hydroxides build one main supplier for additional carbon
%\end{enumerate}

\section{Materials and methods} \label{sec:1}
\subsection{Study site}
The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay. The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system. Extensive wetlands and mangroves dominate this area. The Sin\'{u} river has its origin in the northern part of the western Andean mountains, and between 1938 and 1945 changed its course creating a new delta \citep{suarez2004}. The current estuarine mangrove system was established in the previous river delta. 
The area includes mainly 3 mangrove species. While fringe forests are dominated by the species \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}. \textit{Laguncularia racemosa} occurs in both forest types.

\subsection{Field sampling}
Our sampling focused on a set of existing plots previously established to determine the carbon sequestration potential of the mangroves of Cispat\'a bay \citep{Bolivar2015}. We sampled 10 plots up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter. We sampled five soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4). In addition to the 10 plots sampled across the mangrove area, 3 nearby sand cores at Nisperal coast were collected, and also 3 riverbed cores of the Sin\'{u} river near the city of Monter\'{i}a, which is located 70 km south from Cispat\'{a} bay (Fig. \ref{fig:1}). Generally, cores without layer changes were divided into sections every 20 cm. We divided conspicuous layer changes at the boundary. Because only the general mineralogical and elemental composition of the sand and river samples was important - we considered them as end members - those samples were not divided into segments. Instead, the whole cores down to 1 m were used as individual samples. \par
Additionally, we selected one fringe and one basin mangrove plot (P21 and C4) to measure bulk density. At each site, a soil pit was dug and samples were collected using sampling rings with a volume of 98.52 cm$^3$. We collected four depth levels at plot P21 (0-20, 20-40, 40-60, 60-80 cm) and five depth levels (+ 80-100 cm) at plot C4. Sampling for bulk density was replicated three times for each depth level. In total, we obtained 60 soil samples out of the mangrove area, 6 end-member samples, and 27 samples for bulk density measurements.
\label{sec:2}

\subsection{Laboratory analyses}
After collection, samples were oven-dried at 70$^{\circ}$C for 5 days. Samples for bulk density measurements were oven dried at 105$^{\circ}$C. We calculated soil bulk density as dry mass divided by fresh volume, which matches the used sampling rings \mbox{(V = 98.52 cm$^3$)}. Each sample was ground for 3 min at a frequency of 25 Hz using a Retsch MM 400 ball mill.

%\paragraph{Elemental analysis}
We conducted elemental analyses of percent  carbon (\%TC) and nitrogen (\%TN) in all samples by dry combustion (Vario Max, Elementar Analysensysteme GmbH, Hanau, Germany). Organic carbon was later removed by ignition at 450$^{\circ}$C for 16 hours, and inorganic carbon  (\%IC) was then determined using the same elemental analyzer. Organic carbon concentrations were estimated by subtracting \%IC from \%TC. TOC contents (Mg ha$^{-1}$) were calculated using the obtained bulk densities for basin and fringe mangrove soils multiplied by each plot depth interval (cm) and \%OC.

%\paragraph{$\delta$\textsuperscript{13}C Signatures} 
We used $\updelta^{13}$C values of the sampled material and compared them with the  $^{13}$C/$^{12}$C ratio of the two chosen end-members as indicators for the origin of the carbon. We measured $^{13}$C of all samples using a Finnigan MAT IRMS coupled with an EA 1100 elemental analyzer. Ali-j3 (Acetanilide-Jena3) and Caf-j3 (a caffeine sample from a `Traube synthesis' in large supply) were chosen as internal working standards \citep{werner2001}. All elemental and isotopic analyses were conducted at the Max Planck Institute for Biogeochemistry in Jena, Germany.

%\paragraph{X-ray diffraction analysis}
We conducted X-ray diffraction (XRD) measurements for qualitative and quantitative phase analyses \citep{spiess2009} on 12 representative samples including 4 end-member samples, and 4 samples of basin and fringe mangroves each. Samples were measured 20 min each, from 5 to 60 $^{\circ}2\uptheta$. We determined each mineral phase using the powder diffraction file (PDF) data. We also conducted a Rietveld refinement using Topas (Bruker). To define the type of clays included in the samples, we further measured each before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels. XRD analyses were conducted at the laboratory for Mineralogy and Geochemistry of the Friedrich-Schiller University in Jena, Germany.

%\paragraph{Dithionite- and oxalate extractable Fe and Al} 
We determined iron and aluminum in acid-ammonium-oxalate extracts (pH 3.0) and in sodium-citrate-dithionite extracts (pH 7.3) \citep{Schwertmann, Holmgren1967}. We selected 46 samples and 2 standard soils for these measurements. While sodium dithionite was used to extract both crystalline and amorphous oxides, the oxalate method extracted only amorphous oxides. The actual measurement of crystalline and amorphous iron and aluminium oxides was performed using an atomic emission spectrometer with inductive coupled plasma (ICP-AES, Optima 3300DV, PerkinElmer, Norwalk, USA). These analyses were performed at the SpecLab of the Max Planck Institute for Biogeochemistry.

\subsection{Data analysis}
Comparison of group means for the different measured variables were performed using standard analysis-of-variance procedures. In all cases, we fitted a linear model between the response and the group variable (mangrove type) using the {\tt lm} function in the R language for statistical computing (The R Foundation, Vienna). Anova tables with summary statistics were obtained using the function {\tt anova} in R, including $F$-statistic, degrees of freedom, and $p$-values. Data and code to reproduce all results presented here can be obtained from the following repository \url{https://github.com/crlsierra/mangroveCstabilization.git}.

%\begin{equation}
%TOC(MgC/ha)=OC (\%) \cdot \textit{bulk density (g/cm$^3$)} \cdot \textit{depth interval (cm)}
%\label{equ:1}
%\end{equation}
\section{Results}

We found important differences between fringe and basin mangroves in terms of \%OC ($p$-value $< 0.001$; analysis of variance $F$-test, $F = 54.13$ with 49 degrees of freedom) and \%TN ($p$-value $< 0.001$, $F = 27.1$ with 49 d.f.) (Fig. \ref{fig:2}). In general, \%OC and \%TN decreased with soil depth, with a more clear trend for \%OC than \%TN.

Both end-members, sampled at the Sin\'{u} river in Monter\'{i}a and at Nisperal beach, show \%OC and \%TN concentrations close to zero, which indicates that they include almost no organic matter. In contrast, sand plots displayed the highest \%IC concentration of around 10\% (Fig. \ref{fig:2}). The lack of organic carbon in the fluvial sediments is evidence against our initial hypothesis of carbon imports to the mangrove system through sedimentation.

We obtained much higher bulk densities for basin than for fringe mangrove soil samples. Average bulk density for fringe mangrove soils was 0.16 g cm$^{-3}$, while for basin mangrove soils it was 1.11 g cm$^{-3}$ (Table \ref{tab:1}). These values confirm previous results on the same area obtained by \citet{Bolivar2015}.

Differences in bulk densities between the two forest types resulted in a completely different distribution of C compared to the previous results based on \%OC alone. A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}). The differences in TOC between basin and fringe mangrove types are significant according to the one-way analysis of variance test ($p$-value $= 0.032$, $F = 4.88$, 48 d.f.). TOC decreased with soil depth within the first four (basin) and three (fringe) layers. There was an increase in TOC for the last measured depth intervals of basin and fringe soils, which can be traced back to higher \%OC values of those layers.

We measured a higher proportion of negative $\updelta^{13}$C values for fringe soils \mbox{(-28 to -29 \permil)} than for basin mangroves soils \mbox{(-25 to -29 \permil)} (Fig. \ref{fig:3}).  Depth intervals 20-40 and 40-60 cm of basin soils show outliers with equal $^{13}$C values compared to the Sin\'{u} river end-member samples with values of around \mbox{-25 \permil}. The sand end-member samples have the most positive $^{13}$C values with approximately -3 \permil.  Strong variations in $^{13}$C are conspicuous for the depth intervals 20-40 and 40-60 cm of basin mangrove boxplots. We found no identifiable continuous trend between $^{13}$C values and depth. However, differences in $^{13}$C mean values between fringe and basin mangroves were statistically significant ($p$-value $< 0.001$, $F = 24.35$, 49 d.f.).

There were no differences between basin and fringe mangroves in terms of their composition of the mineral fraction (Table \ref{tab:2}). The proportion of mineral soil is bigger in basin than in fringe mangroves. Fringe mangrove soils show in contrast a higher proportion of halite than basin soils.

The XRD pattern of sand samples measured from 5 to 60 $^{\circ}2\uptheta$ provided characteristic $^{\circ}2\uptheta$ intensities for the phases aragonite, calcite and quartz. The XRD measurement combined with a Rietveld refinement yielded a mineralogical composition of 95\% aragonite, 4\% calcite and 1\% quartz.
River samples showed high intensities for quartz and sodium feldspar (albite) components, as well as peaks for the clay minerals illite and clinochlore. The semi-quantitative distribution calculated by PDF data indicated that quartz is the main representative of the mineral fraction of the Sin\'{u} river soils (Fig. \ref{fig:4}).

XRD patterns for both mangrove soils, but in particular for basin soils, showed a similar mineralogical distribution compared to the Sin\'{u} river sediments (Fig. \ref{fig:5}). XRD patterns in fringe soils showed high intensities for halite, which was also found in basin mangrove soils, but not in a comparable distribution ratio.
XRD measurements conducted on the extracted clay mineral fraction yielded peaks for the mineral phases clinochlore, illite, quartz and albite. PDF data additionally identified corundum, which is related to the ceramic panel surface composition. We found major proportions for clinochlore and quartz in the clay size fraction.

The oxalate extraction dissolved much of the iron and aluminum from the amorphous materials, whereas the dithionite extraction dissolved only little of the crystalline iron oxides as well as little of the amorphous materials \mbox{(Fig. \ref{fig:6})}. Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $< 0.001$, $F = 16.35$, 39 d.f.). Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $F = 33.45$, 39 d.f.).
Basin soil samples had two times higher concentrations of Al (Al$_d$+Al$_o$) than fringe soils, and five times higher Fe values. While riverine samples had similar Al and Fe concentrations than those of basin mangroves in all 4 extraction patterns, metal contents of sand samples were constantly low. Fe$_o$ contents differed extremely for basin soils, especially for the first 2 depth intervals. Both basin and fringe soils showed an increase of metal oxides within the first depth intervals and in turn a decrease within the deeper layers.

\section{Discussion}
Our results confirmed previously observed differences in \%OC, bulk density and \%TOC between the basin and the fringe mangrove soils \citep{Bolivar2015}.  Furthermore, our measurements of carbon isotopes and mineralogy helped us to establish the potential origin of the carbon stored in both mangrove types. In the following we will discuss these difference and the potential implications of our findings.

\subsection{Differences in carbon storage}
Lower values of bulk density in fringe mangrove soils can be explained by their high organic matter content (wood residues, leaf debris and roots). Roots claim a large proportion of the soil volume, which strongly decreases bulk density. Instead, basin mangrove samples of Cispat\'{a} bay are characterized by a dense silty composition including a small organic part, which results in higher bulk densities. The increase in bulk density with depth in both mangrove types is likely the result of hydrostatic pressure and time inside the mangrove forest belowground.\par
The higher proportion of organic matter in fringe than in basin soils also leads to higher \%OC values for fringe soils. Basin mangrove soils instead, are characterized by a small portion of organic topsoil. Because \%TN concentrations correlate with \%OC, they also show lower values for basin mangroves. That both end-members (river sediments and sand) are mostly consisting of mineral components is illustrated by their \%OC contents of nearly 0 \%. Because the \%IC concentration of 10 \% of sand samples is similar to the used pure calcium carbonate standard, we concluded that the sand end member is mostly composed of carbonates.

The decrease of \%OC with soil depth is likely the result of the interaction between decomposition, vertical transport of organic matter, and leaching of dissolved carbon in water. Because soil microorganisms utilize nitrogen and bacteria fix nitrogen, the concentration of \%TN also decreases with depth. The fact that concentrations of \%OC and \%TN decrease continuously with soil depth, but still show layers with higher concentrations in depths of 80-100 cm (basin) and 60-80 cm (fringe), may reflect differences in sedimentation rates over time \citep{Bolivar2015}. In Cispat\'{a} bay, silting processes linked to changes in the position of the Sin\'{u} river delta, current sea level rise, flooding regime and fluvial inputs, can generate deep organic layers that may cause the increase of \%OC with depth for both mangrove types \citep{suarez2004}.

Higher TOC values in basin mangrove soils reflect higher rates of organic matter accumulation. According to  \citet{Bolivar2015}, the percentage of clay is similar between both mangrove types. However, the silt fraction dominates in all soil profiles in basin mangroves, while sand dominates in fringe mangrove soils. It has been well established that soil particles with greater surface area, as typical of finer textures like those found in basin mangroves, deteriorate drainage conditions which in turn increase retention of organic matter \citep{prasad2008}. Because TOC contents are linked to \%OC, we found also an increase of TOC in depth levels of 80-100 cm (basin) and 60-80 cm (fringe).

Our results confirm previous studies that found important differences in TOC between fringe and basin mangrove soils \citep{Bolivar2015}. Furthermore, basin mangrove soils showed a significantly higher range of in-situ produced carbon than fringe mangrove soils, and with it a higher carbon storage of 661$\pm$116 MgC ha$^{-1}$ (0-100 cm) compared to 320$\pm$60 MgC ha$^{-1}$ (0-80 cm). Based on the \%IC concentration results, we infer that marine sediments across Cispat\'{a} bay have no influence on additional carbon entering via tidal flooding, which is supported by the lack of \%IC content in the analyzed mangrove soil samples compared to the sands of Nisperal beach.

\subsection{Origin of carbon}
$\updelta^{13}$C values of fringe mangrove sediments were more $^{13}$C depleted than basin sediments, which is a strong indication that this carbon is more plant derived. C3 plants fractionate $^{13}$C during photosynthesis, with values from \mbox{-22 to -38 $\permil$}, while C4 plants show values between \mbox{-8 to -15 $\permil$} \citep{Farquhar1989}. \textit{R. mangle}, a typical C3 species, mostly occurs in the fringe area of Cispat\'{a} bay \citep{Bolivar2015} and is therefore the main contributor of $^{13}$C in this type of forest. In terms of its origin, this carbon is very likely produced in situ in the area. Basin mangrove sediments instead, show a wider range of $\updelta^{13}$C values and more positive values. This could mean either that the basin area has a higher contribution by C4 plants (e.g. grasses), or that there is some contribution from mineral sources. Because the most common basin mangrove species, \textit{A. germinans}, is also a C3 plant and there is likely little contribution by C4 grasses, we assumed that basin mangrove sediments have some influence by deposited sediments from the Sin\'{u} river. This assumption is also supported by the fact that basin samples tend to have more enriched $^{13}$C values, close to those found for the Sin\'{u} river samples (-25 $\permil$). According to \citet{ruttenberg1997} and \citet{powers2005}, $\updelta^{13}$C values of tropical mineral soils range from -23 to -26 $\permil$, which would underpin mineral derived $^{13}$C values in Cispat\'{a} bay. Because differences in $^{13}$C values were significant between basin and fringe sediments, and because fringe samples only have plant derived $^{13}$C, we conclude that the fringe area is not influenced by the Sin\'{u} river delta. It also does not show any influence through marine sediments, because it does not have any enriched values comparable to the sand end member (-5 $\permil$).

The $\updelta^{13}$C analyses confirmed that additional carbon present in basin soils is only terrestrial and not marine derived. Moreover, it shows that additional terrestrial carbon only influences basin mangrove sediments, while carbon in fringe soils is exclusively produced in situ by the plants.\par

Using XRD, we found that the mineralogical composition of the sediments in Cispat\'{a} bay was similar to the composition of the Sin\'{u} river sediments in Monter\'{i}a. We therefore conclude that basin and fringe mangrove sediments were transported by the river and have their distant origin in that river basin. Quartz, albite and clay minerals are the major components of those sediments. Additionally, plot R5 shows intense peaks for the mineral halite in depths between 80-90 cm. 
Because halite crystallizes by the evaporation of sea water and intense salinization \citep{McCaffrey}, its occurrence verifies that the fringe mangrove area has tidal influence of hyper saline sea water. Furthermore, halite confirms the sediment air exposition, since it needs evaporation to be formed. That halite reaches that intense proportion in depths of 80-90 cm can be attributed to the changing course of the Sin\'{u} river. According to \citet{suarez2004}, the river course passed the northern main land of Cispat\'{a} bay between 1849 and 1938, where plot R5 is located. Accordingly, the sedimentation rate increased in that period. However, before that period, this region was more influenced by saline sea water, which results in intense halite peaks. That basin soils also contain halite, shows that they were also occasionally  flooded. \par

%Depending on their cation-exchange capacity, clay minerals are able to bind metal oxides. Because cation-exchange sites on clay and humic colloids have negative charge, cations (metals) from the soil solution must satisfy this charge, so that mineral and organic surfaces appear to be charged neutral. Cations are retained on these clay sites by electrostatic forces. At the same time, Al and Fe oxide interactions with clays are pH dependent (variable charges). At low pH, where the oxides carry sufficient positive charge, they precipitate on clay surfaces. These coatings, once formed, are stable at higher pHs. Precipitation of oxides at high pH occurs as phases separated from the clays \citep{goldberg1989}. According to \citet{goldberg1989}, observed precipitation of oxides showed that only amorphous Fe-precipitates were obtained with the clay substrate. However, in the case of quartz, precipitated oxides had increased crystallinity, particularly for the samples with high iron concentrations. Transferred to our study sites, these previous results help to explain that the increased proportion of metal oxides in basin soils goes together with a more mineralogical composed soil. \par

Along with the $^{13}$C results, the mineralogical analyses also confirmed that basin mangrove sediments are terrestrially derived. This finding is also reflected in higher metal oxide and hydroxide contents in basin than in fringe mangrove soils. Concentrations of iron and aluminum oxides in basin mangroves were similar to those of the Sin\'{u} river samples, which is a strong indication of fluvial deposition of inland sediments.

%The contents of metal oxides in Cispat\'{a} bay are generally low compared to other soils \citep{mckeague1966}. Also, mangrove soils in Cispat\'{a} bay do not show typical reddish colour changes, which are an indication for the presence of iron oxides (especially hematite). A continuous grey colouring of all non-humus layers shows further the only presence of clay minerals \citep{segalen1971}. Only plot C1 contains yellow and beige coloured layers in depths of 40-80 cm, which indicates the occurrence of goethite ($\upalpha$-Fe$^{3+}$O(OH)). Beside goethite, also gibbsite (Al(OH)$_3$), indicated by a white colour, is a typical hydroxide for mangrove sediments \citep{souza2008} and soils of warm and humid areas of the world \citep{segalen1971}. Al and Fe are mostly present as amorphous products, e.g. as alumino- or ferrogels. The negligible small amounts of metal oxides and hydroxides in all 3 sand samples indicate furthermore, that those transported as well as in situ produced oxides get fixed in mangrove soils. An increase of Al and Fe is apparent within the first depth intervals, but there is no continuos trend with depth. Again, this shows the connection between mineral soil horizons and Al and Fe bonding and the smaller amount of Al and Fe in humus layers. \par

That mineral surfaces of Al and Fe oxides and hydroxides adsorb dissolved organic matter (DOM) has been well established \citep{tipping1981, Oades1988, kaiser2000, Mikutta2006}. It is therefore very likely that metal oxides and hydroxides bind and preserve C in basin mangrove soils. According to \citet{kaiser2000}, the capacity to adsorb DOM relates to the presence of Al and Fe oxides and hydroxides. The sorption of DOM derived from decomposition to Al and Fe oxyhydroxides involves strong complexation bondings between surface metals and acidic organic ligands, particularly with those associated with aromatic structures. The strength of the sorption relates further to the surface properties of the sorbing mineral phase. \citet{kaiser2000} found that dissolved organic matter sorption is strongly enhanced by hydrous oxide coatings and particularly by amorphous Al(OH)$_3$, which indicates that amorphous hydroxides bind C in basin soils of Cispat\'{a} bay. \citet{tipping1981} describes moreover, that the extent of adsorption of DOM increases with decreasing pH. Because mangrove soils at the Colombian Caribbean coast have an acid character both at \textit {A. germinans} and \textit {R. mangle} forests \citep{urrego2014}, a stronger C binding onto oxides and hydroxides is substantiated in this region.

All together, we found no evidence for our initial hypothesis that C in mangroves of Cispat\'a bay, particularly in basin soils, have a large contribution from sediments transported by the Sin\'u river. Instead, we found that the sediments do play an important role for stabilizing in situ produced carbon, but mostly by providing mineral surfaces for C binding.

\subsection{Implications}
Our results provide strong evidence for an important role of sediment mineralogy, particularly iron and aluminum oxides, in providing mineral surfaces for the adsorption of dissolved carbon and long-term C retention \citep{Oades1988} in mangrove soils . These results add a new dimension to the more traditional studies of carbon origin in mangrove soils where the source of the carbon is considered either marine or terrestrially derived \citep{lacerda, Bouillion}, without considering the role of sediments in dissolved carbon retention. This mechanism may play a large role in explaining observed spatial variability in carbon storage \citep{Alongi2012, Jardine2014}. Also, since retention of dissolved carbon in mangrove soils can reduce rates of carbon exports to the ocean \citep{Adame2011}, mineral surfaces of sediments may provide a large potential for carbon sequestration in mangrove ecosystems located around river deltas. According to our results, carbon storage can be twice as high in mangroves with dense aggregation of minerals than in more organic mangrove soils, and therefore it is very relevant to explore this carbon sequestration mechanism in other delta regions of the world.

\section{Conclusions}
Based on analyses of carbon concentration, bulk density, stable isotopes of carbon, and mineralogy, we found that most carbon stored in soils of Cispat\'a Bay, Colombia, is produced in situ, with little evidence of carbon imported to the area either by fluvial or marine sedimentation. Interior basin mangroves store significantly more carbon in soils than the more ocean exposed fringe mangroves. Sediments transported by the Sin\'u river and deposited in the delta region contain negligible amounts of organic carbon, but the mineralogical composition of these sediments favors the adsorption of dissolved carbon on charged mineral surfaces, which explains the larger levels of C storage in this type of mangroves.

Our study highlights the importance of fluvial sediment transport in providing a substrate for carbon stabilization through mineral protection. This mechanism for soil carbon storage has been little studied previously in mangrove ecosystems, but it has large implications for determining their potential for long-term carbon sequestration.

%\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
% figure caption is below the figure
%
% For tables use

\begin{acknowledgements}
We would like to thank Prof. Dr. Juraj Majzlan for providing mentoring and access to the XRD lab at the Friedrich-Schiller-Universit\"at Jena. Thanks to INVEMAR Colombia, particularly Paula C. Sierra, for facilitating logistics and access to the research site. Financial support was provided by the Alexander von Humboldt Foundation through the International Climate Protection Fellowship, and the Max Planck Institute for Biogeochemistry.
\end{acknowledgements}

\clearpage

\section*{Figures}

% Fig 1
\begin{figure}[H]
\centering
%  \includegraphics[width=9cm]{Figures/Kartefern.jpg}
% figure caption is below the figure
\caption{Sampling plots in Cispat\'{a} Bay and Monter\'{i}a, Colombia, ArcGIS-source: http://www.diva-gis.org/gdata; accessed on June 07, 2016; fringe plots: P21, R1, R4, R5, R6; basin plots: P16, C1, C2, C3, C4; end-members: Sand Nisperal, Sin\'{u} riverbed}
\label{fig:1}       % Give a unique label
\end{figure}

% Fig 2
\begin{figure}[H]
%\centering
%  \includegraphics[width=12cm]{Figures/Fig2.pdf}
% figure caption is below the figure
\caption{Observed values of percent inorganic carbon (\%IC) for both mangrove types and end members; and percent organic carbon (\%OC) and percent nitrogen (\%TN) by sampling depth aggregated across plots. }
\label{fig:2}       % Give a unique label
\end{figure}

% Fig 3
\begin{figure}[H]
%\centering
%  \includegraphics[width=12cm]{Figures/Fig3.pdf}
% figure caption is below the figure
\caption{Measured $\updelta^{13}$C in a) mangrove types and end members, b) basin mangroves by depth, and c) fringe mangroves by depth. }
\label{fig:3}       % Give a unique label
\end{figure}

% Fig 4
\begin{figure}[H]
%\centering
%  \includegraphics[width=12cm]{Figures/EndMembers.pdf}
% figure caption is below the figure
\caption{Two exemplary XRD patterns ($\uplambda$ = 1.5406 \AA) of both measured end-members: Nisperal beach and Sin\'{u} river; figures also include semi-quantitative distribution of detected minerals}
\label{fig:4}       % Give a unique label
\end{figure}

% Fig 5
\begin{figure}[H]
%\centering
%  \includegraphics[width=12cm]{Figures/R5_C3.pdf}
% figure caption is below the figure
\caption{Two exemplary XRD patterns ($\uplambda$ = 1.5406 \AA) for the fringe (left) and basin (right) mangrove soils; figures also include semi-quantitative distribution of detected minerals}
\label{fig:5}       % Give a unique label
\end{figure}

% Fig 6
\begin{figure}[H]
%\centering
%  \includegraphics[width=12cm]{Figures/Fig6.pdf}
% figure caption is below the figure
\caption{Oxalate (o)- and dithionite (d) extracted metal oxides for each mangrove type and soil depth.}
\label{fig:6}       % Give a unique label
\end{figure}

\clearpage

\section*{Tables}

%Table 1
\begin{table}[H]
% table caption is above the table
\caption{Percent organic carbon, bulk density and total organic carbon by mangrove type and soil depth. Values in parentheses indicate standard deviation}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lllll}
\hline\noalign{\smallskip}
Mangrove type & Depth [cm] & OC [\%]  & Bulk density [g/cm$^3$] & TOC [MgC/ha]\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Basin & 0-20 & 14.63 (10.09) & 1.01 (0.06) & 295.61 (203.77)\\
 & 20-40 & 7.12 (7.19) & 1.25 (0.10) & 161.99 (147.07)\\
 & 40-60 & 3.22 (1.93) & 1.06 (0.19) & 68.35 (40.95)\\
 & 60-80 & 2.96 (1.39) & 1.10 (0.16) & 65.21 (30.58)\\
 & 80-100 & 4.15 (2.24) & 0.84 (0.08) & 69.78 (37.60)\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Fringe & 0-20 & 31.32 (2.82) & 0.16 (0.01) & 100.24 (9.03)\\
 & 20-40 & 23.61 (8.70) & 0.13 (0.02) & 62.79 (22.56)\\
 & 40-60 & 16.12 (10.92) & 0.16 (0.05) & 51.57 (34.94)\\
 & 60-80 & 29.29 (0.55) & 0.18 (0.01) & 105.44 (1.99)\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

% Table 2
\begin{table}[H]
% table caption is above the table
\caption{General mineralogical composition of sediments in Cispat\'{a} Bay as measured by XRD analyses.}
\label{tab:2}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
Class & Mineral  & Formula\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Silicates & Clinochlore & (Mg,Fe$^{2+}$)$_5$Al(Si$_3$Al)O$_{10}$(OH)$_8$\\
& Illite & (K,H$_3$O)Al$_2$(Si$_3$Al)O$_{10}$(H$_2$O,OH)$_2$\\
& Albite & NaAlSi$_3$O$_8$\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Oxides/hydrox. & Quartz & SiO$_2$\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Carbonates & Aragonite & CaCO$_3$\\
& Calcite & CaCO$_3$\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Halides & Halite & NaCl\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

% BibTeX users please use one of
%\bibliographystyle{spbasic}   % basic style, author-year citations
%\bibliography{Reference}

   %\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
%\begin{thebibliography}{}
\begin{thebibliography}{24}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{Adame and Lovelock(2011)}]{Adame2011}
Adame MF, Lovelock CE (2011) Carbon and nutrient exchange of mangrove forests
  with the coastal ocean. Hydrobiologia 663(1):23--50,
  \doi{10.1007/s10750-010-0554-7},
  \urlprefix\url{https://doi.org/10.1007/s10750-010-0554-7}

\bibitem[{Alongi(2012)}]{Alongi2012}
Alongi DM (2012) Carbon sequestration in mangrove forests. Carbon management
  3(3):313--322

\bibitem[{Bol{\'\i}var(2015)}]{Bolivar2015}
Bol{\'\i}var J (2015) {Soil and root carbon in two types of mangroves in
  Cispat\'a bay (Colombian Caribbean).} Master's thesis, Universidad Nacional
  de Colombia sede Medell\'in

\bibitem[{Bouillon et~al(2008)Bouillon, Connolly, and Lee}]{Bouillion}
Bouillon S, Connolly R, Lee S (2008) Organic matter exchange and cycling in
  mangrove ecosystems: Recent insights from stable isotope studies. Journal of
  Sea Research 59(1):44 -- 58,
  \doi{https://doi.org/10.1016/j.seares.2007.05.001},
  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S1385110107000469},
  mangrove Macrobenthos Special Issue

\bibitem[{Donato et~al(2011)Donato, Kauffman, Murdiyarso, Kurnianto, Stidham,
  and Kanninen}]{Donato2011}
Donato DC, Kauffman JB, Murdiyarso D, Kurnianto S, Stidham M, Kanninen M (2011)
  Mangroves among the most carbon-rich forests in the tropics. Nature Geosci
  4(5):293--297, \urlprefix\url{http://dx.doi.org/10.1038/ngeo1123}

\bibitem[{Farquhar et~al(1989)Farquhar, Ehleringer, , and
  Hubick}]{Farquhar1989}
Farquhar GD, Ehleringer JR, , Hubick KT (1989) Carbon isotope discrimination
  and photosynthesis. Annual Review of Plant Physiology and Plant Molecular
  Biology 40(1):503--537, \doi{10.1146/annurev.pp.40.060189.002443}

\bibitem[{Holmgren(1967)}]{Holmgren1967}
Holmgren GGS (1967) A rapid citrate-dithionite extractable iron procedure1.
  Soil Science Society of America Journal 31(2):210--211,
  \urlprefix\url{http://dx.doi.org/10.2136/sssaj1967.03615995003100020020x}

\bibitem[{Jardine and Siikam\"aki(2014)}]{Jardine2014}
Jardine SL, Siikam\"aki JV (2014) A global predictive model of carbon in
  mangrove soils. Environmental Research Letters 9(10):104,013

\bibitem[{Kaiser and Guggenberger(2000)}]{kaiser2000}
Kaiser K, Guggenberger G (2000) {The role of DOM sorption to mineral surfaces
  in the preservation of organic matter in soils}. Organic Geochemistry
  31(7):711--725

\bibitem[{Lacerda et~al(1995)Lacerda, Ittekkot, and Patchineelam}]{lacerda}
Lacerda L, Ittekkot V, Patchineelam S (1995) Biogeochemistry of mangrove soil
  organic matter: a comparison betweenrhizophoraandavicenniasoils in
  south-eastern brazil. Estuarine, Coastal and Shelf Science 40(6):713 -- 720,
  \doi{https://doi.org/10.1006/ecss.1995.0048},
  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0272771485700480}

\bibitem[{McCaffrey et~al(1987)McCaffrey, Lazar, and Holland}]{McCaffrey}
McCaffrey MA, Lazar B, Holland HD (1987) The evaporation path of seawater and
  the coprecipitation of br (super -) and k (super +) with halite. Journal of
  Sedimentary Research 57(5):928--937,
  \doi{10.1306/212F8CAB-2B24-11D7-8648000102C1865D},
  \urlprefix\url{http://jsedres.geoscienceworld.org/content/57/5/928},
  \eprint{http://jsedres.geoscienceworld.org/content/57/5/928.full.pdf}

\bibitem[{Mikutta et~al(2006)Mikutta, Kleber, Torn, and Jahn}]{Mikutta2006}
Mikutta R, Kleber M, Torn MS, Jahn R (2006) Stabilization of soil organic
  matter: Association with minerals or chemical recalcitrance? Biogeochemistry
  77(1):25--56, \doi{10.1007/s10533-005-0712-6},
  \urlprefix\url{https://doi.org/10.1007/s10533-005-0712-6}

\bibitem[{Oades(1988)}]{Oades1988}
Oades JM (1988) The retention of organic matter in soils. Biogeochemistry
  5(1):35--70, \doi{10.1007/BF02180317},
  \urlprefix\url{https://doi.org/10.1007/BF02180317}

\bibitem[{Powers and Veldkamp(2005)}]{powers2005}
Powers JS, Veldkamp E (2005) {Regional variation in soil carbon and $\delta$13C
  in forests and pastures of northeastern Costa Rica}. Biogeochemistry
  72(3):315--336

\bibitem[{Prasad and Ramanathan(2008)}]{prasad2008}
Prasad MBK, Ramanathan A (2008) {Sedimentary nutrient dynamics in a tropical
  estuarine mangrove ecosystem}. {Estuarine, Coastal and Shelf Science}
  80(1):60--66

\bibitem[{Ruttenberg and Goni(1997)}]{ruttenberg1997}
Ruttenberg K, Goni M (1997) {Phosphorus distribution, C: N: P ratios, and
  $\delta$ 13 C oc in arctic, temperate, and tropical coastal sediments: tools
  for characterizing bulk sedimentary organic matter}. Marine Geology
  139(1):123--145

\bibitem[{Schwertmann(1964)}]{Schwertmann}
Schwertmann U (1964) Differenzierung der eisenoxide des bodens durch extraktion
  mit ammoniumoxalat-l{\"o}sung. Zeitschrift f{\"u}r Pflanzenern{\"a}hrung,
  D{\"u}ngung, Bodenkunde 105(3):194--202, \doi{10.1002/jpln.3591050303},
  \urlprefix\url{http://dx.doi.org/10.1002/jpln.3591050303}

\bibitem[{Serrano(2004)}]{suarez2004}
Serrano BE (2004) {The Sin{\'u} river delta on the northwestern Caribbean coast
  of Colombia: bay infilling associated with delta development}. Journal of
  South American Earth Sciences 16(7):623--631

\bibitem[{Spie{\ss} et~al(2009)Spie{\ss}, Teichert, Schwarzer, Behnken, and
  Genzel}]{spiess2009}
Spie{\ss} L, Teichert G, Schwarzer R, Behnken H, Genzel C (2009) {Moderne
  R{\"o}ntgenbeugung: R{\"o}ntgendiffraktometrie f{\"u}r
  Materialwissenschaftler, Physiker und Chemiker}. Springer-Verlag

\bibitem[{Spohn and Giani(2012)}]{Spohn2012}
Spohn M, Giani L (2012) {Carbohydrates, carbon and nitrogen in soils of a
  marine and a brackish marsh as influenced by inundation frequency}.
  Estuarine, Coastal and Shelf Science 107:89--96

\bibitem[{Spohn et~al(2013)Spohn, Babka, and Giani}]{Spohn2013}
Spohn M, Babka B, Giani L (2013) {Changes in soil organic matter quality during
  sea-influenced marsh soil development at the North Sea coast}. Catena
  107:110--117

\bibitem[{Tipping(1981)}]{tipping1981}
Tipping E (1981) {The adsorption of aquatic humic substances by iron oxides}.
  {Geochimica et Cosmochimica Acta} 45(2):191--199

\bibitem[{Urrego et~al(2014)Urrego, Molina, and Su{\'a}rez}]{urrego2014}
Urrego L, Molina E, Su{\'a}rez J (2014) {Environmental and anthropogenic
  influences on the distribution, structure, and floristic composition of
  mangrove forests of the Gulf of Urab{\'a} (Colombian Caribbean)}. {Aquatic
  Botany} 114:42--49

\bibitem[{Werner and Brand(2001)}]{werner2001}
Werner RA, Brand WA (2001) {Referencing strategies and techniques in stable
  isotope ratio analysis}. Rapid Communications in Mass Spectrometry
  15(7):501--519

\end{thebibliography}

%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
%\bibitem{RefJ}
% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
%\end{thebibliography}

\end{document}
% end of file template.tex

### 2017-11-13 Code_and_data/reprod.R
Fixed directory structure and updated metadata information

@@ -1,13 +1,13 @@
 ## Read the data
-setwd("~/Repos/Mangrove/Mineralogy/Measurements/")
+setwd("~/Repos/mangroveCstabilization/Code_and_Data/")
 Ergebnisse<-read.csv("Ergebnisse_vereinfacht.csv", stringsAsFactors = FALSE)
 C_N_13C<-read.table("C_N_13C.csv", sep=";", dec=".", header = TRUE)
-TOC<-read.csv("Bulk_density/TOC.csv")
-Al_Fe<-read.csv2("Al_Fe/Al_Fe.csv", dec=".")
-Sand_3<-read.csv("XRD/Sand_3.csv")
-Sinu_3<-read.csv("XRD/Sinu_3.csv")
-R5_80_90<-read.csv("XRD/R5_80-90.csv")
-C3_80_85<-read.csv("XRD/C3_80-85.csv")
+TOC<-read.csv("TOC.csv")
+Al_Fe<-read.csv2("Al_Fe.csv", dec=".")
+Sand_3<-read.csv("Sand_3.csv")
+Sinu_3<-read.csv("Sinu_3.csv")
+R5_80_90<-read.csv("R5_80-90.csv")
+C3_80_85<-read.csv("C3_80-85.csv")
 
 
 #######################################################
@@ -33,7 +33,7 @@ anova(lm(Fe_Sum~Type, data=Al_Fe[1:41,]))
 
 #######################################################
 # Fig 2
-pdf("~/Repos/Mangrove/MA_Paper/Springer_Template/Figures/Fig2.pdf")
+#pdf("../Manuscript/Figures/Fig2.pdf")
 par(mfrow=c(3,1))
 boxplot(Cinorg~Type, data=Ergebnisse,  par(mar=c(4,4,1,1)), ylab="Inorganic carbon (%)")
 
@@ -46,10 +46,10 @@ boxplot(Ntotal~Depth_adjusted, data=Ergebnisse[1:51,], par(mar=c(5,4,1,1)), xaxt
 axis(1, at=1:10, labels=c("0-20","20-40","40-60","60-80","80-100","0-20","20-40","40-60","60-80","80-100"), las=1)
 legend("topright", c("basin", "fringe"), col = c("peachpuff","lightsteelblue"), bg = "gray90", pch=15, bty="n", y.intersp = 0.7)
 par(mfrow=c(1,1))
-dev.off()
+#dev.off()
 
[…]

### 2017-12-12 Manuscript/carbonStabilizationMangroves.tex
Fixed issue with LaTeX preamble
Minor changes based on pdf provided by reviewer 2

@@
 % your LaTeX will extract the file if required
-\begin{filecontents*}{example.eps}
+%\begin{filecontents*}{example.eps}
 %!PS-Adobe-3.0 EPSF-3.0
@@
 %%EndComments
-gsave
-newpath
-  20 20 moveto
-  20 220 lineto
-  220 220 lineto
-  220 20 lineto
-closepath
-2 setlinewidth
-gsave
-  .4 setgray fill
-grestore
-stroke
-grestore
-\end{filecontents*}
-%
-\RequirePackage{fix-cm}
+%gsave
+%newpath
+%  20 20 moveto
+%  20 220 lineto
+%  220 220 lineto
+%  220 20 lineto
+%closepath
+%2 setlinewidth
+%gsave
+%  .4 setgray fill
+%grestore
+%stroke
+%grestore
+%\end{filecontents*}
+%%
+%\RequirePackage{fix-cm}
 %
@@
 \label{intro}
-Although mangroves are ecosystems with some of the largest levels of carbon storage on earth \citep{Donato2011, Alongi2012}, there are  large variations on the amount of C stored in various systems, particularly for soils.
+Although mangroves are ecosystems with some of the largest levels of carbon storage on earth \citep{Donato2011, Alongi2012}, there are  large variations on the amount of C stored in various systems, particularly soils.
 For instance, mangrove ecosystems dominated by \emph{Rhizophora} spp.
@@
 The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay.
-The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system.
+The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system that covers approximately 5,098 ha.
 Extensive wetlands and mangroves dominate this area.
@@
 The current estuarine mangrove system was established in the previous river delta.
-The area includes mainly 3 mangrove species.
-While fringe forests are dominated by the species \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}.
-\textit{Laguncularia racemosa} occurs in both forest types.
+Fringe forests are dominated by  \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}. \textit{Laguncularia racemosa} occurs in both forest types.
 
@@
 Samples for bulk density measurements were oven dried at 105$^{\circ}$C.
-We calculated soil bulk density as dry mass divided by fresh volume, which matches the used sampling rings \mbox{(V = 98.52 cm$^3$)}.
+We calculated soil bulk density as dry mass divided by fresh volume \mbox{(V = 98.52 cm$^3$)}.
 Each sample was ground for 3 min at a frequency of 25 Hz using a Retsch MM 400 ball mill.
@@
 There were no differences between basin and fringe mangroves in terms of their composition of the mineral fraction (Table \ref{tab:2}).
-The proportion of mineral soil is bigger in basin than in fringe mangroves.
+The proportion of mineral soil is larger in basin than in fringe mangroves.
 Fringe mangrove soils show in contrast a higher proportion of halite than basin soils.
@@
 
-The oxalate extraction dissolved much of the iron and aluminum from the amorphous materials, whereas the dithionite extraction dissolved only little of the crystalline iron oxides as well as little of the amorphous materials \mbox{(Fig.
+The oxalate extraction dissolved much of the poorly crystalline Fe and Al oxides from the amorphous materials, whereas the dithionite extraction dissolved the crystalline Fe oxides as well as the amorphous materials \mbox{(Fig.
 \ref{fig:6})}.
@@
 However, the silt fraction dominates in all soil profiles in basin mangroves, while sand dominates in fringe mangrove soils.
-It has been well established that soil particles with greater surface area, as typical of finer textures like those found in basin mangroves, deteriorate drainage conditions which in turn increase retention of organic matter \citep{prasad2008}.
+It has been well established that soil particles with greater surface area, as typical of finer textures like those found in basin mangroves, decrease drainage and decomposition of organic matter \citep{prasad2008}.
 Because TOC contents are linked to \%OC, we found also an increase of TOC in depth levels of 80-100 cm (basin) and 60-80 cm (fringe).

### 2018-06-27 Reviews/Response.tex
Started working on second round of revisions

\documentclass[11pt]{bgcletter}
\usepackage{hyperref}

\name{Dr.\ Carlos A. Sierra}
\signature{ \vspace{-2cm} Carlos A. Sierra, PhD}
\email{csierra}
\telephone{6133}
\begin{document}
\begin{letter}{Dr. Takashi Asaeda\\
 Editor in Chief \\ Wetlands Ecology and Management}

\opening{Dear Dr. Aseda}
Thank you very much for your consideration of our manuscript and for the opportunity to submit a revised version. We made several changes to the manuscript based on the reviewers' comments. More prominently, we added two new sections, one describing the statistical methods used in the study, and another discussing the implications of our findings. For reproducibility and clarity in our data analysis methods, we also now provide the data and the code necessary to reproduce all results and figures.

 In the text below, we provide answers ({\color{blue} blue font}) to all reviewers' comments ({\it italics font}).

{\bf Reviewer 1} \\
{\it The manuscript of Volkel et al. strengthens previous findings showing that most soil carbon in mangroves is from local production. This is one of the few manuscripts I have seen for this region in Latin America, so it is regionally important. The conclusions on soil carbon origin were supported by mineralogical analyses and stable isotopes, although the stable isotope analyses was incorrectly performed. The design of the study is well conducted, although the number of samples is relatively low (n= 10 plots). The manuscript is well written. See below my detailed suggestions:}

{\color{blue} We thank the reviewer for recognizing the value of the manuscript. Although the reviewer mentions that this study has weaknesses in terms of the isotopic analysis conducted and its representativeness, we believe this is a misunderstanding and clarify these issues in the more specific comments below. }

\begin{itemize}
\item {\it L9- Is better to extrapolate to 1m, instead of 0.8m, to facility comparisons among studies (see (Kauffman and Donato, 2012)}

{\color{blue} We can easily report the values up to 1 m for basin mangroves, but for fringe mangroves soil depth does not go below 0.8 m. It would be wrong to make such an extrapolation. We therefore provide in the abstract a value up to 1 m for the basin mangroves and only up to 0.8 cm for fringe mangroves.}

\item {\it L107- Brand of EL analyser?}
{\color{blue} Elementar. Information added to the manuscript. }

\item {\it L114- Isotopes are ``values'', not signatures, they are not constant. You have to include in your end member analyses: mangrove material, riverine material, marine material (phytoplankton?). You could plot them as a triangle or analyse with mixing model. See (Fry, 2013)}

{\color{blue} We changed `signatures' to `values' as suggested by the reviewer. Please note that we used reported values for the mangrove material as end member in our analysis. We acknowledge that the isotopic analysis could be improved if more end members were available; however, we also want to point out that we not only used C isotopes to identify source contributions, but also mineralogy analyses of river and marine sediments.}

%However, at this point we cannot add a third end member to the analysis since the sampling campaign did not include mangrove material. We considered the sand samples as end members for the marine material, which help to identify inorganic contributions. We acknowledge that the isotopic analysis could be improved if more end members were available; however, we also want to point out that we not only used C isotopes to identify source contributions, but also mineralogy analyses of river and marine sediments. }

\item {\it L117- where were the samples analysed?}

{\color{blue}At the Max Planck Institute for Biogeochemistry in Jena, Germany. This information was added to the manuscript.}

\item {\it L119- What are the standards used for the isotope analyses? }

{\color{blue} The standards were Ali-j3 and Caf-j3. We added a reference were these standards are reported.}

\item {\it L134- How were the samples analysed? }

{\color{blue} We added a section on data analysis describing basic statistical analyses used in this manuscript.}

\item {\it Results \\
You need to add F, df, and p values to all your comparisons. }

{\color{blue} Done.}

\item {\it L156- Make a section on statistical analyses within the Methodology. }

{\color{blue} Done. We are also providing as supplementary material all code and data to reproduce the statistical analyses and figures reported.}

\item {\it L166- Stable isotope values are ratios compared to a standard, so it is obvious that if the value is close to zero, it means that is similar to the standard. No need to explain it. }

{\color{blue} Sentence removed as suggested.}

\item {\it L172- What about sediment deposited in the basin forest after hurricanes, storm surges? Would that contribute to higher riverine sediment in the basin forest? }

{\color{blue} This is irrelevant for our site. During the entire historical hurricane record, no hurricane have arrived to the Colombian Caribbean coast (see Figure below). This is due to the geographical location of this coast, which is protected by the northern part of the South American continent from the main tracks of Atlantic hurricanes. }

\includegraphics[scale=0.15]{noaa-hurricane-tracker.jpg}

\item {\it Fig 2- It would be more interested to have the mean values for the fringe vs basin vs sand vs river material, instead of values for each of the plots. }

{\color{blue} Figure changed as requested.}

\item {\it Fig 3- Better to call the end members `riverine' instead of `Sinu'. Need to include other end members to the analyses such as mangrove material and phytoplankton (maybe some published values ?)}

{\color{blue} We made a number of changes in this figure to improve clarity, including the issue suggested by the reviewer. Published values for other end members are provided in the text.}

\item {\it L221- This is not really interesting or relevant to this paper}

{\color{blue} Although it may not be very interesting, we consider it is relevant to report that the measured values for the sand end member are similar to those of the standard.}

\item {\it L223- And decomposition}

{\color{blue} We changed `heterotrophic consumption by microorganisms' by `decomposition' as suggested.}

\item {\it L259- $\delta$13C is not produced by anything, is just a ratio between the heavy (13C) and the light (12C) isotope. }

{\color{blue} We change `producer' by `contributor' to address this comment.}

\item {\it L266- Need to include all the relevant end members}

{\color{blue} As mentioned before, we used three end-members in our analysis. We also acknowledge that end members for phytoplankton are missing, but our analysis is complemented by the analysis of mineral composition of the sediments. }

\item {\it L357- I am not sure this is the conclusion of the manuscript. Although most of the C seems to be derived from mangroves, basin forests seem to have a higher contribution of other source, either riverine or marine? You could use the N:C values to clarify sources. }

{\color{blue} We found that soils in basin mangroves are dominated by riverine sediments; however, carbon contents in the river samples were close to zero. Instead, we found similar mineralogy among both sediments, but much larger C storage in the basin mangroves. It is well stablished that C has a large affinity to sorb in mineral surfaces of sediments, and our analysis is indicative that the in-situ derived carbon has been stabilized in this mineral surfaces. We believe this is strong evidence to reach this conclusion. }

\item {\it L367- I did not understand how the mineralogical composition facilitates dissolved carbon absorption. Regardless, it doesn't seem to matter much if most of the material is from mangrove origin as it has been shown in previous studies}

{\color{blue} The mineral composition of the sediments determines the surface area available for dissolved organic molecules to sorb. This is well known for different soils and mineral surfaces. To make this more clear, we added a number of additional references and explain this process with more detail in the text. }

\end{itemize}

{\bf Reviewer 2} \\
{\it This paper aims to study source of carbon in a mangrove-wetland complex in Colombia by exploring sediment characteristics. Although a potentially important and interesting study, it appears authors/researchers were not able to gather evidence that might be sufficient to draw the conclusions they presented. This reviewer believes that after critical revision, adding significant new information and providing additional analysis/interpretation their work could be a good addition to the literature, but in its present form, it is not fit for publication.}

{\color{blue} We thank the reviewer for recognizing the importance of this work as well as providing critical comments. We acknowledge that some information was missing in the first version of the manuscript, and we made an effort to include missing details.}

{\it This MS could benefit immensely with a clear layout of sampling plan - sampling stations/plots, mangrove types, distribution and general habitat characteristics, elaborating the logic behind the selection of sampling plots. Attributes such as water depth, tidal range, salinity, proximity to the river mouth, fresh water discharge (rainfall patterns) and tidal mixing etc should be provided. Size of estuarine complex, and annual rainfall etc can be included in study area description.}

{\color{blue} An important detail that we forgot to mention in the submitted version is that this study was part of a much larger project in which a number of plots were stablished previously to estimate carbon stocks in the area. We did not selected the plots ourselves, but rather sampled the set of existing plots. Criteria for plot selection are described in other publications and we refer to them in the new version of the manuscript. }

{\it Also, Figure 1 (map of study location) indicates that river Sinu flows almost 10-12 km west of the sampling region. There is not clear suggestion in the figure or text that samples sites receive waters from River Sinu. This reviewer is somewhat perplexed and unable to understand the hypothesis and choice of the study design. Perhaps some clarification and effort in explaining this via text and figures would be beneficial for the readers.}

{\color{blue} We acknowledge that this point may not have been well explained in the area description section of the previous version of the manuscript. The Sin\'u River changed its course in 1938, and before this year it discharged in the study area. This is well documented in Serrano (2004), and we provide now a better explanation in the description of the study site.}

{\it Some important information is missing in the laboratory analysis section. For this entire section there is no reference/no citation. This reviewer encourages authors to cite appropriate methods that were followed for analysis, and include all necessary and important details.}

{\color{blue} We added necessary citations regarding the different methods used.}

{\it For calculating carbon storages in the sampled sites, bulk density of only one fringe site (P21) and one basin site (C4) was used for all plots. There is a large amount of spatial variability in soils even at small resolution and it is important to use site specific bulk density to calculate site specific storages. This reviewer has a serious reservation against this methodology.}

{\color{blue} Please note the extreme differences in bulk density between the two mangrove types reported in Table 1. While basin mangroves have a bulk density of around 1.1, fringe mangroves are around 0.1 g cm$^{-3}$. We obtained similar results in a previous study that included more replicates (Bolivar 2015), and the numbers reported here simply confirm previous measurements. It is unlikely that these differences are due to spatial variability within mangrove types. In fact, these differences in bulk density are an inherent characteristic of the different mangrove types. The fringe mangroves are stablished in the old Sin\'u river delta sediments, while the fringe mangroves advance the cost line and establish over coastal sands. }

{\it This methods section does not include any description how data was analyzed. No statistical information is provided. What programs were used. Whether any tests for normalcy were performed? It is a standard practice to include such information in the MS. }

{\color{blue} We added a new section describing the statistical procedures and the software used. In addition, we provide the code and data to reproduce all results.}

{\it This reviewer believes that the discussion section of this MS should be revised with more critical inputs, highlighting the new addition to this field. This reviewer commends authors for tedious field work and many hours of data entry and subsequent analysis, but it becomes difficult to see value of this research unless a more rigorous and through discussion is presented. The researchers could also report the broader implications of their research… i.e. How this understanding would be beneficial and for whom.}

{\color{blue} Thanks for the recommendation. We added a paragraph to the discussion about the broader implications of this study.}

{\it There are multiple incidences of discrepancies in sentence syntax and complex sentence formulation that diminishes readability and clarity of the text. This reviewer has suggested moving around text, and rewriting sentences to impart clarity. Such corrections will improve paper's readability and should be relatively straight-forward to fix. }

{\color{blue} Thanks for pointing this out. We edited the text in several parts to reduce sentence complexity and improve clarity.}

\vspace{2em}
{\bf Comments by reviewer 2 on pdf version of the article} \\
The pdf version of the article provided by the reviewer contained a large number of highlighted text, comments, and strikethrough text. We provided below answers to the comments, and the strikethrough text was removed from the manuscript. However we did not performed any action on the highlighted text since there was no indication by the reviewer on the reason the text was highlighted.

{\color{blue} Can you provide an approximate size of this estuarine lagoon complex?} \\
{\it The total area is 5,098 ha. We added this value to the site description section. }

{\color{blue} how about Crystalline iron oxides? } \\
{\it The oxalate extraction dissolves mostly poorly crystalline iron oxides, while the dithionite extraction removes the crystalline Fe oxides. We added this to the text. }

{\color{blue} Figure 4, About 85\% is Aragonite. That also clearly suggests CaCO3 presence. }
{\it Yes, this is correct and we point it out in the main text as well. }

{\color{blue} Ln. 240. Consider rewriting. Doesn't convey meaning clearly} \\
{\it Sentence rewritten as suggested.}

{\color{blue} Ln 303. Looks like there is something missing here. You can provide more information here to build up an explanation for your observation. } \\
{\it This sentence was removed from the latest version of the manuscript because it does not add useful information for the message we want to convey. }

\vspace{2em}
We hope this new version adequately addresses reviewer's comments and it is now suitable for publication.

\closing{Sincerely, \\
 \includegraphics[scale=0.7]{../../../../Documents/Personal/firma.jpg}
 }
 \end{letter}

 \end{document}

### 2018-06-27 Manuscript/carbonStabilizationMangrovesV2.tex
Started working on second round of revisions

--- 2017-12-12 Manuscript/carbonStabilizationMangroves.tex
@@
 \subsection{Study site}
-The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay.
+The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay (Fig.
+\ref{fig:1}).
 The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system that covers approximately 5,098 ha.
@@
 Our sampling focused on a set of existing plots previously established to determine the carbon sequestration potential of the mangroves of Cispat\'a bay \citep{Bolivar2015}.
-We sampled 10 plots up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter.
-We sampled five soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4).
+We sampled 10 plots of 500 m$^2$ up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter (Eijkelkamp bi-partite gouge auger 04.03, Giesbeek, The Netherlands).
+We selected five randomly chosen points within each plot to extract soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4).
 In addition to the 10 plots sampled across the mangrove area, 3 nearby sand cores at Nisperal coast were collected, and also 3 riverbed cores of the Sin\'{u} river near the city of Monter\'{i}a, which is located 70 km south from Cispat\'{a} bay (Fig.
 \ref{fig:1}).
-Generally, cores without layer changes were divided into sections every 20 cm.
-We divided conspicuous layer changes at the boundary.
-Because only the general mineralogical and elemental composition of the sand and river samples was important - we considered them as end members - those samples were not divided into segments.
-Instead, the whole cores down to 1 m were used as individual samples.
-\par
+Generally, cores without layer changes were divided into sections every 20 cm to explore differences of carbon concentration with depth.
+We divided conspicuous layer changes at the boundary, except for the sand and river sediment samples that were analyzed as one single sample for mineralogical and elemental composition.
+
 Additionally, we selected one fringe and one basin mangrove plot (P21 and C4) to measure bulk density.
-At each site, a soil pit was dug and samples were collected using sampling rings with a volume of 98.52 cm$^3$.
+At each site, a soil pit was dug three meters away from one of the corners of the plot and samples were collected using sampling rings with a volume of 98.52 cm$^3$.
 We collected four depth levels at plot P21 (0-20, 20-40, 40-60, 60-80 cm) and five depth levels (+ 80-100 cm) at plot C4.
@@
 We calculated soil bulk density as dry mass divided by fresh volume \mbox{(V = 98.52 cm$^3$)}.
-Each sample was ground for 3 min at a frequency of 25 Hz using a Retsch MM 400 ball mill.
+Each sample was ground for 3 min at a frequency of 25 Hz using a ball mill (Retsch MM 400, Haan, Germany).
 
@@
 %\paragraph{$\delta$\textsuperscript{13}C Signatures} 
-We used $\updelta^{13}$C values of the sampled material and compared them with the  $^{13}$C/$^{12}$C ratio of the two chosen end-members as indicators for the origin of the carbon.
+We used $\updelta^{13}$C values of the sampled material and compared them with the  $^{13}$C/$^{12}$C ratio of the two chosen end-members as indicators for the origin of the carbon \citep{Fry2006}.
 We measured $^{13}$C of all samples using a Finnigan MAT IRMS coupled with an EA 1100 elemental analyzer.
@@
 We determined each mineral phase using the powder diffraction file (PDF) data.
-We also conducted a Rietveld refinement using Topas (Bruker).
-To define the type of clays included in the samples, we further measured each before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels.
+We also conducted a Rietveld refinement using the Topas software for XRD analysis (Bruker Corporation).
+To define the type of clays included in the samples, we further measured each sample before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels.
 XRD analyses were conducted at the laboratory for Mineralogy and Geochemistry of the Friedrich-Schiller University in Jena, Germany.
@@
 We obtained much higher bulk densities for basin than for fringe mangrove soil samples.
-Average bulk density for fringe mangrove soils was 0.16 g cm$^{-3}$, while for basin mangrove soils it was 1.11 g cm$^{-3}$ (Table \ref{tab:1}).
+Average $\pm$ standard deviation of bulk density for fringe mangrove soils was 0.15 $\pm$ 0.02 g cm$^{-3}$, while for basin mangrove soils it was 1.07 $\pm$ 0.12 g cm$^{-3}$ (Table \ref{tab:1}).
 These values confirm previous results on the same area obtained by \citet{Bolivar2015}.
 
-Differences in bulk densities between the two forest types resulted in a completely different distribution of C compared to the previous results based on \%OC alone.
+Differences in bulk densities between the two forest types resulted in a completely different distribution of C compared to the previous results based on \%OC alone (Table \ref{tab:1}).
 A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}).
@@
 Lower values of bulk density in fringe mangrove soils can be explained by their high organic matter content (wood residues, leaf debris and roots).
-Roots claim a large proportion of the soil volume, which strongly decreases bulk density.
+Roots claim a large proportion of the soil volume, which strongly decreases bulk density \citep{Bolivar2015}.
 Instead, basin mangrove samples of Cispat\'{a} bay are characterized by a dense silty composition including a small organic part, which results in higher bulk densities.
-The increase in bulk density with depth in both mangrove types is likely the result of hydrostatic pressure and time inside the mangrove forest belowground.\par
+%The increase in bulk density with depth in both mangrove types is likely the result of hydrostatic pressure and time inside the mangrove forest belowground.\par
+
 The higher proportion of organic matter in fringe than in basin soils also leads to higher \%OC values for fringe soils.
@@
 
-The decrease of \%OC with soil depth is likely the result of the interaction between decomposition, vertical transport of organic matter, and leaching of dissolved carbon in water.
+The decrease of \%OC with soil depth is likely the result of the interaction between decomposition, vertical transport of organic matter, and leaching of dissolved carbon in water \citep{Elzein1995, Braakhekke2013, Mathieu2015}.
 Because soil microorganisms utilize nitrogen and bacteria fix nitrogen, the concentration of \%TN also decreases with depth.
@@
 
-Along with the $^{13}$C results, the mineralogical analyses also confirmed that basin mangrove sediments are terrestrially derived.
+Along with the $^{13}$C results, the mineralogical analyses also confirmed that basin mangrove sediments are terrestrially derived as opposed to marine derived.
 This finding is also reflected in higher metal oxide and hydroxide contents in basin than in fringe mangrove soils.
@@
 \section{Conclusions}
-Based on analyses of carbon concentration, bulk density, stable isotopes of carbon, and mineralogy, we found that most carbon stored in soils of Cispat\'a Bay, Colombia, is produced in situ, with little evidence of carbon imported to the area either by fluvial or marine sedimentation.
+Based on analyses of carbon concentration, stable isotopes of carbon, and mineralogy, we found that most carbon stored in soils of Cispat\'a Bay, Colombia, is produced in situ, with little evidence of carbon imported to the area either by fluvial or marine sedimentation.
 Interior basin mangroves store significantly more carbon in soils than the more ocean exposed fringe mangroves.
@@
 % BibTeX users please use one of
-%\bibliographystyle{spbasic}   % basic style, author-year citations
-%\bibliography{Reference}
+\bibliographystyle{spbasic}   % basic style, author-year citations
+\bibliography{Reference}
 
@@
 % Non-BibTeX users please use
-%\begin{thebibliography}{}
-\begin{thebibliography}{24}
-\providecommand{\natexlab}[1]{#1}
-\providecommand{\url}[1]{{#1}}
-\providecommand{\urlprefix}{URL }
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem[{Adame and Lovelock(2011)}]{Adame2011}
-Adame MF, Lovelock CE (2011) Carbon and nutrient exchange of mangrove forests
-  with the coastal ocean. Hydrobiologia 663(1):23--50,
-  \doi{10.1007/s10750-010-0554-7},
-  \urlprefix\url{https://doi.org/10.1007/s10750-010-0554-7}
-
-\bibitem[{Alongi(2012)}]{Alongi2012}
-Alongi DM (2012) Carbon sequestration in mangrove forests. Carbon management
-  3(3):313--322
-
-\bibitem[{Bol{\'\i}var(2015)}]{Bolivar2015}
-Bol{\'\i}var J (2015) {Soil and root carbon in two types of mangroves in
-  Cispat\'a bay (Colombian Caribbean).} Master's thesis, Universidad Nacional
-  de Colombia sede Medell\'in
-
-\bibitem[{Bouillon et~al(2008)Bouillon, Connolly, and Lee}]{Bouillion}
-Bouillon S, Connolly R, Lee S (2008) Organic matter exchange and cycling in
-  mangrove ecosystems: Recent insights from stable isotope studies. Journal of
-  Sea Research 59(1):44 -- 58,
-  \doi{https://doi.org/10.1016/j.seares.2007.05.001},
-  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S1385110107000469},
-  mangrove Macrobenthos Special Issue
-
-\bibitem[{Donato et~al(2011)Donato, Kauffman, Murdiyarso, Kurnianto, Stidham,
-  and Kanninen}]{Donato2011}
-Donato DC, Kauffman JB, Murdiyarso D, Kurnianto S, Stidham M, Kanninen M (2011)
-  Mangroves among the most carbon-rich forests in the tropics. Nature Geosci
-  4(5):293--297, \urlprefix\url{http://dx.doi.org/10.1038/ngeo1123}
-
-\bibitem[{Farquhar et~al(1989)Farquhar, Ehleringer, , and
-  Hubick}]{Farquhar1989}
-Farquhar GD, Ehleringer JR, , Hubick KT (1989) Carbon isotope discrimination
-  and photosynthesis. Annual Review of Plant Physiology and Plant Molecular
-  Biology 40(1):503--537, \doi{10.1146/annurev.pp.40.060189.002443}
-
-\bibitem[{Holmgren(1967)}]{Holmgren1967}
-Holmgren GGS (1967) A rapid citrate-dithionite extractable iron procedure1.
-  Soil Science Society of America Journal 31(2):210--211,
-  \urlprefix\url{http://dx.doi.org/10.2136/sssaj1967.03615995003100020020x}
-
-\bibitem[{Jardine and Siikam\"aki(2014)}]{Jardine2014}
-Jardine SL, Siikam\"aki JV (2014) A global predictive model of carbon in
-  mangrove soils. Environmental Research Letters 9(10):104,013
-
-\bibitem[{Kaiser and Guggenberger(2000)}]{kaiser2000}
-Kaiser K, Guggenberger G (2000) {The role of DOM sorption to mineral surfaces
-  in the preservation of organic matter in soils}. Organic Geochemistry
-  31(7):711--725
-
-\bibitem[{Lacerda et~al(1995)Lacerda, Ittekkot, and Patchineelam}]{lacerda}
-Lacerda L, Ittekkot V, Patchineelam S (1995) Biogeochemistry of mangrove soil
-  organic matter: a comparison betweenrhizophoraandavicenniasoils in
-  south-eastern brazil. Estuarine, Coastal and Shelf Science 40(6):713 -- 720,
-  \doi{https://doi.org/10.1006/ecss.1995.0048},
-  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0272771485700480}
-
-\bibitem[{McCaffrey et~al(1987)McCaffrey, Lazar, and Holland}]{McCaffrey}
-McCaffrey MA, Lazar B, Holland HD (1987) The evaporation path of seawater and
-  the coprecipitation of br (super -) and k (super +) with halite. Journal of
-  Sedimentary Research 57(5):928--937,
-  \doi{10.1306/212F8CAB-2B24-11D7-8648000102C1865D},
-  \urlprefix\url{http://jsedres.geoscienceworld.org/content/57/5/928},
-  \eprint{http://jsedres.geoscienceworld.org/content/57/5/928.full.pdf}
-
-\bibitem[{Mikutta et~al(2006)Mikutta, Kleber, Torn, and Jahn}]{Mikutta2006}
-Mikutta R, Kleber M, Torn MS, Jahn R (2006) Stabilization of soil organic
-  matter: Association with minerals or chemical recalcitrance? Biogeochemistry
-  77(1):25--56, \doi{10.1007/s10533-005-0712-6},
-  \urlprefix\url{https://doi.org/10.1007/s10533-005-0712-6}
-
-\bibitem[{Oades(1988)}]{Oades1988}
-Oades JM (1988) The retention of organic matter in soils. Biogeochemistry
-  5(1):35--70, \doi{10.1007/BF02180317},
-  \urlprefix\url{https://doi.org/10.1007/BF02180317}
-
-\bibitem[{Powers and Veldkamp(2005)}]{powers2005}
-Powers JS, Veldkamp E (2005) {Regional variation in soil carbon and $\delta$13C
-  in forests and pastures of northeastern Costa Rica}. Biogeochemistry
-  72(3):315--336
-
-\bibitem[{Prasad and Ramanathan(2008)}]{prasad2008}
-Prasad MBK, Ramanathan A (2008) {Sedimentary nutrient dynamics in a tropical
-  estuarine mangrove ecosystem}. {Estuarine, Coastal and Shelf Science}
-  80(1):60--66
-
-\bibitem[{Ruttenberg and Goni(1997)}]{ruttenberg1997}
-Ruttenberg K, Goni M (1997) {Phosphorus distribution, C: N: P ratios, and
-  $\delta$ 13 C oc in arctic, temperate, and tropical coastal sediments: tools
-  for characterizing bulk sedimentary organic matter}. Marine Geology
-  139(1):123--145
-
-\bibitem[{Schwertmann(1964)}]{Schwertmann}
-Schwertmann U (1964) Differenzierung der eisenoxide des bodens durch extraktion
-  mit ammoniumoxalat-l{\"o}sung. Zeitschrift f{\"u}r Pflanzenern{\"a}hrung,
-  D{\"u}ngung, Bodenkunde 105(3):194--202, \doi{10.1002/jpln.3591050303},
-  \urlprefix\url{http://dx.doi.org/10.1002/jpln.3591050303}
-
-\bibitem[{Serrano(2004)}]{suarez2004}
-Serrano BE (2004) {The Sin{\'u} river delta on the northwestern Caribbean coast
-  of Colombia: bay infilling associated with delta development}. Journal of
-  South American Earth Sciences 16(7):623--631
-
-\bibitem[{Spie{\ss} et~al(2009)Spie{\ss}, Teichert, Schwarzer, Behnken, and
-  Genzel}]{spiess2009}
-Spie{\ss} L, Teichert G, Schwarzer R, Behnken H, Genzel C (2009) {Moderne
-  R{\"o}ntgenbeugung: R{\"o}ntgendiffraktometrie f{\"u}r
-  Materialwissenschaftler, Physiker und Chemiker}. Springer-Verlag
-
-\bibitem[{Spohn and Giani(2012)}]{Spohn2012}
-Spohn M, Giani L (2012) {Carbohydrates, carbon and nitrogen in soils of a
-  marine and a brackish marsh as influenced by inundation frequency}.
-  Estuarine, Coastal and Shelf Science 107:89--96
-
-\bibitem[{Spohn et~al(2013)Spohn, Babka, and Giani}]{Spohn2013}
-Spohn M, Babka B, Giani L (2013) {Changes in soil organic matter quality during
-  sea-influenced marsh soil development at the North Sea coast}. Catena
-  107:110--117
-
-\bibitem[{Tipping(1981)}]{tipping1981}
-Tipping E (1981) {The adsorption of aquatic humic substances by iron oxides}.
-  {Geochimica et Cosmochimica Acta} 45(2):191--199
-
-\bibitem[{Urrego et~al(2014)Urrego, Molina, and Su{\'a}rez}]{urrego2014}
-Urrego L, Molina E, Su{\'a}rez J (2014) {Environmental and anthropogenic
-  influences on the distribution, structure, and floristic composition of
-  mangrove forests of the Gulf of Urab{\'a} (Colombian Caribbean)}. {Aquatic
-  Botany} 114:42--49
-
-\bibitem[{Werner and Brand(2001)}]{werner2001}
-Werner RA, Brand WA (2001) {Referencing strategies and techniques in stable
-  isotope ratio analysis}. Rapid Communications in Mass Spectrometry
-  15(7):501--519
-
-\end{thebibliography}
 

### 2018-06-27 Reviews/Response2.tex
Started working on second round of revisions

\documentclass[11pt]{bgcletter}
\usepackage{hyperref}

\name{Dr.\ Carlos A. Sierra}
\signature{ \vspace{-2cm} Carlos A. Sierra, PhD}
\email{csierra}
\telephone{6133}
\begin{document}
\begin{letter}{Dr. Takashi Asaeda\\
 Editor in Chief \\ Wetlands Ecology and Management}

\opening{Dear Dr. Aseda}
Thank you very much for your consideration of our manuscript and for the opportunity to submit a revised version. We made several changes to the manuscript based on the reviewers' comments. More prominently, we \dots

 In the text below, we provide answers ({\color{blue} blue font}) to all reviewers' comments ({\it italics font}).

{\bf Reviewer 4} \\

{\it I recognise the considerable efforts the authors have taken to undertake this study and complete a thorough laboratory analysis of the samples collected. Moreover, I also appreciate the necessity of understanding sediment dynamics and carbon accumulation with respect to mangrove systems, which this study addresses. Overall, what is lacking to me here in this manuscript is a sufficient level of detail in the general presentation and a recognition of the limitations of the study. This extends to a consistent failure to validly support various statements made throughout the manuscript by making appropriate reference to suitable literature sources. Furthermore, the detailed interpretations and contrasts between individual habitat `zones', which have been inferred from the basic statistical analyses used, are somewhat na\"ive.  The discussions section contains some very confidently-worded interpretations from the results, but does not provide sufficient information to support these
(be it either from literature sources or logical argument).  Finally, the authors appear to assume, throughout the introduction, methods and results section, that the reader will understand the rationale behind the field sampling and analytical methods used. Links between the hypotheses and why these methods were used are not given until well into the discussion section. The reader is left to try to figure this out for themselves.}

\begin{enumerate}
\item {\it Ln. 37 Describe what habitat or environmental characteristics make up a `basin', or `fringe' mangrove. Again - a map of the area would be very helpful here! (Figure 1 is not mentioned in the text).}

{\color{blue} Habitat and environmental characteristics of the basing and fringe mangroves are described in the Mathrials and Methods section. Here in the introduction, we are describing the main motivation of the study and therefore it is not the most appropriate place to describe the study site. Figure 1 is mentioned in the text in the Methods section.}

\item {\it Ln. 60-62 You need to provide information on why this is the case - and what your rationale for this method is. I presume - that 13C (CO2) is preferentially taken up by mangrove trees during photosynthesis, therefore accounting for higher ratios in the mangrove sediments? But you haven't given any sort of explanation for this. Your rationale does not seem to explain why allochthonous organic carbon sediments (derived from elsewhere in this catchment would not also be enriched with 13C). What is the rationale supporting this line of inquiry - and most importantly - where are the literature sources to validate it?}

\item {\it Ln 71-73 A map showing the location might be helpful. (Figure 1 is not mentioned in the manuscript text)}

{\color{blue} Figure 1 is a map that shows the location of the study site and the location of the plots. It was initially mentioned in section 2.2, but we added an earlier reference in section 2.1 to address this comment.}

\item {\it Ln. 77 How? Why?}

{\color{blue} The cause of the change of the Sin\'u river course is not well known, however it is well documented in Serrano (2004), which we cite in the manuscript for further information.}

\item {\it Ln 86 Using what apparatus? What diameter? What depth? How were the sample positions determined? How were your sample locations randomised? (This is a FUNDAMENTAL basic requirement for the parametric statistics you have used).}

{\color{blue} The diameter and depth were already given in the sentence above. We added a description of the soil corer and information on the random selection of points. We know this is fundamental, and quite obvious, so for this reason we didn't include it in the previous version, but we add it now to address this comment.}

\item {\it Ln 87-88 A map to show the layout of these plots? How big are they? How far are they spaced apart from each other? How were their original positions determined? What do the designations P, R and C actually refer to and why is it necessary to use them?}

{\color{blue} The location of the plots is presented in Figure 1. The size of the plots is 500 m$^2$. We added this information to the manuscript. Plot designations are relevant for consistency with previously published manuscripts form the same site. }

\item {\it Ln 91-92 Why was this done? What was the rationale for choosing 20cm divisions? Where is a citation to support this?}

{\color{blue} This was done to see if we can detect differences by soil depth instead of analyzing the entire soil core. This is commonly done in soil science and usually does not require any specific justification.}

\item {\it Ln 93-96 I can't follow these two sentences. What is an 'end member'? Who or what were 'them'?}

{\color{blue} This sentence was reworded for clarity.}

\item {\it Ln 98 Where in the plot? How was this position determined? Was this at a position you simply liked the look of?}

{\color{blue} Actually outside the plot to avoid effects on the trees inside the plot. The location was three meters away from a randomly selected corner of the plot. }

\item {\it Ln 101-102 So was a (single) soil pit dug? (Ln. 98) Or were 3 soil pits dug? If these replicates came from a single soil pit - how can they be considered replicates?}

{\color{blue} No, two soil pits in plots P21 and C4. This is explicitly mentioned in the manuscript.}

\item {\it ln 105-107 Why were these protocols chosen? (Citations?)}

{\color{blue} These are standard soil processing protocols.}

\item {\it Ln 108 Manufacturer name and location.}

{\color{blue} Information add to text.}

\item {\it Ln 118 -120 Do you have a reference to support this inference and further detail to support why it is valid to assume that this method allows for indication of the 'origin of the carbon'?}

{\color{blue} This is basic knowledge on isotope source partitioning analysis. We assumed most reads are a familiar with this concept, however we added a reference to a textbook to satisfy reviewer curiosity. }

\item {\it Ln 131 What is this refinement method, why did you use it? Where is the justification in the literature for using it in this way?}

{\color{blue} This is simply the name of the technique. See \url{https://en.wikipedia.org/wiki/Rietveld_refinement}}

\item {\it Ln 131 ``Topas'' Is this a citation, a manufacturer name, a piece of software, or an alternative name for the piece of apparatus?}

{\color{blue} Topas is a software and Brucker the company who produces it. We added more information to the sentence for clarity.}

\item {\it Ln 132 Measured what of each? The clay? The mass of the sample? The Rietveld / Topas number?}

{\color{blue} Each sample.}

\item {\it Ln 136 - 139 I miss a rationale, here or in the introduction section, for why these analyses were necessary. What were you hoping to achieve by doing this? What were your hypotheses?}

{\color{blue} The rationale is presented in the introduction, where we mention that we use analyses of the structure and composition of the clay minerals to help in determining the origin of the stored carbon.}

\item {\it Ln 157 I remain more than a little confused by your analysis and the conclusions you have made from them. Your statistical analysis (an ANOVA, examining the variance in a range of response variables according to 4 categorical 'zones' within this habitat system: basin, fringe, river, sand) does not allow you to determine the statistical significance of differences between individual habitat zones, even if these contrasts appear (visually) in the figures. You will need to include a Tukey post-hoc HSD analysis of means to be able to validly make such inferences on a pair-wise basis.}

{\color{blue} Sorry for the confusion. In the ANOVA we didn't compare four categorical zones as the reviewer thought, but only the two groups of mangrove types: fringe and basin. Therefore the Tukey post-hoc test is not needed. }

\item {\it Ln 169-170 Error terms for these mean values??}

{\color{blue} Standard deviation values were added. }

\item {\it Ln 172-173 How can you make such a definitive inference, based on your observations alone? How do you know that X resulted in Y? (The reader is not convinced - as you haven't provided any explanation why!)}

{\color{blue} This is inferred by simply multiplying bulk density by \% organic carbon. Also, you can reach this conclusion by simply looking at the results of Table 1. We do not think we need to give a complex argumentation for this result since it is almost trivial. However, in the new version we point the reader to Table 1 to see this in more detail. }

\item {\it Ln 175 -177 I do not see how you can assume to have directly assessed such a contrast, based on the description of the statistical methods and analyses you have used.}

{\color{blue} Again, we are not comparing the four groups here, only the two mangrove types, and for this reason it is possible to express the result in the way we do in this paragraph.}

\item {\it Ln 188-190 The statistical test you have described in your methods section did not test for direct differences between your individual habitat zone categories, yet a statistically significant contrast between two individual categories is precisely what you are stating here. }

{\color{blue} i.d.e.m.}

\item Ln 229-230 But you haven't at any point described how or why it did this.

\item {\it Ln 236 Reference to support this claim? Surely you were sampling root-free soils? Or did your soil samples also contain live root biomass?}

{\color{blue} Reference added. }

\item {\it Ln 239 Wouldn't an increase in bulk density be expected with any soil? Why is this relevant to mangroves? Why is reporting this finding relevant?}

{\color{blue} Yes, this happens in most soils, but in mangroves bulk densities increase strongly due to hydrostatic pressure. However, we agree in that this sentence is not very relevant for the discussion on differences in bulk density among the two mangrove types.}

\item {\it Ln 251-254 On what basis? Do you have a references to support these rationales? It is quite a large assumption otherwise!}

{\color{blue} This is relatively well known for vertical carbon transport in soils. We added a few references to support this claim. }

\item Ln 281 - 288 Some, if not all of the information in this paragraph needs to be in the introduction / methods section. Without such validating information, until now, what you have done has remained pretty much meaningless to the reader.

\item {\it Ln 285 What kind of (plant) species - a sunflower? Wheat? Mangrove tree species?}

{\color{blue} We do not understand this comment by the reviewer. Line 285 explicitly says Rizophora mangle.}

\item {\it Ln 286 - 387 'and is *likely* to be' 
You haven't conducted sufficient analyses to be certain!
Here - and elsewhere - there are far too many 'definitive' statements like this in the manuscript, which are either unverifiable the your analyses you have conducted, or unsupported by relevant literature sources. Both these requirements should be the basic minimum for any peer-reviewed scientific study.}

{\color{blue} We beg to differ. In this paragraph we do not make any definitive statements as stated by the reviewer. When we use a word such as `likely', we are actually implying that this might be the case, but there is uncertainty. Notice that we use wording such as `this assumption is supported', `this could mean either ... or ...' , 'which would underpin', etc. These are all hedged statements that show our interpretation of the results and our favoring for a specific interpretation, but away from a definite true.}

\item {\it Ln 326 As opposed to being from a marine source? This (implied?) contrast should be made more clear.}

{\color{blue} The sentence was completed to make the contrast clear as suggested.}

\item {\it Ln 372 I am perplexed as to how you claim your analyses of bulk density were able to show this. Changes in BD were co-correlated with root density and organic carbon content, that's all. They cannot give any indication of the source of carbon in a soil.}

{\color{blue} Sure, bulk density has nothing to do with a determination of the origin of the carbon. Since this was the opening statement of the conclusions, we simply wanted to list all the analyses that were performed in the study. However, we removed bulk density from this paragraph to avoid confusion.}

\end{enumerate}

\newpage
{\bf Reviewer 5} \\

{\it The manuscript of Volkel et al. aims to investigate the origin of organic carbon in mangrove ecosystem in Colombia using stable isotopes and mineralogical analyses. The authors improve the manuscript following the suggestions of the reviewers. As the previous reviewers have already marked this article is of interest regionally but also as a contribution in a wider contest. I think that in this form the manuscript is almost ready for publication. I have just few comments.}

{\it Reviewer 2 highlighted as in the manuscript a statistical section on methodology was missing. The authors add a chapter on statistical methods without explain if before the ANOVA analysis they have checked the population for normality. This is an important point since a normal distribution is an assumption for ANOVA and many soil characters have not a normal distribution. In the file with your codes, I didn't find any code that is usually used to test for normality.}

{\it Why did you not add data on N/C ratio (or at least C/N ratio) as the reviewer 1 has suggested? You already have values on C and N, adding N/C ration should be a further indicator to support your results and conclusions.}

\vspace{2em}
We hope this new version adequately addresses reviewer's comments and it is now suitable for publication.

\closing{Sincerely, \\
 \includegraphics[scale=0.7]{../../../Documents/Personal/firma.jpg}
 }
 \end{letter}

 \end{document}

### 2018-06-27 Code_and_data/reprod.R

@@ -199,14 +199,21 @@ dev.off()
 
 
 # Mean values with propagated uncertainty
+# For TOC
 B=c(295.61,161.99,68.35,65.21,69.78)
 sum(B)
 sdB=c(203.77,147.07,40.95,30.58,37.60)
 sqrt(sum(sdB^2)/length(sdB))
 
-
 Fr=c(100.24,62.79,51.57,105.44)
 sdF=c(9.03,22.56,51.57,105.44)
 sum(Fr)
 sqrt(sum(sdF^2)/length(sdF))
+
+# For Bulk density
+round(mean(TOC[TOC[,"Type"]=="Basin","BulkDens"]), 2)
+round(sd(TOC[TOC[,"Type"]=="Basin","BulkDens"]), 2)
+
+round(mean(TOC[TOC[,"Type"]=="Fringe","BulkDens"]), 2)
+round(sd(TOC[TOC[,"Type"]=="Fringe","BulkDens"]), 2)
 

### 2018-06-28 Manuscript/carbonStabilizationMangrovesV2.tex
Addressed comments form reviewer 5

@@
 \subsection{Data analysis}
-Comparison of group means for the different measured variables were performed using standard analysis-of-variance procedures.
+We performed statistical tests to determine differences in organic carbon concentrations, total nitrogen, $\delta^{13}$C, Fe- and Al-oxides, among the fringe and basin mangroves.
+After an initial test for normality (Shapiro-Wilk test), we found little evidence to support the normality assumption required in common tools such as ANOVA or t-tests.
+We therefore, performed the non-parametric Mann-Whitney U test, which is equivalent to the two-sample Wilcoxon test \citep{Hollander2015} under the null hypothesis that the samples from the fringe and basin mangroves  differ by a shift location $\mu =0$, and the alternative hypothesis that they different by some other shift location.
+In other words, the test helps to determine whether the samples from the two different mangrove types belong to the same statistical distribution as stated in the null hypothesis.
+Data and code to reproduce all results presented here can be obtained from the following repository \url{https://github.com/crlsierra/mangroveCstabilization.git}.
+
+%Comparison of group means for the different measured variables were performed using standard analysis-of-variance procedures.
 In all cases, we fitted a linear model between the response and the group variable (mangrove type) using the {\tt lm} function in the R language for statistical computing (The R Foundation, Vienna).
@@
 
-We found important differences between fringe and basin mangroves in terms of \%OC ($p$-value $< 0.001$; analysis of variance $F$-test, $F = 54.13$ with 49 degrees of freedom) and \%TN ($p$-value $< 0.001$, $F = 27.1$ with 49 d.f.) (Fig.
+We found important differences between fringe and basin mangroves in terms of \%OC ($p$-value $< 0.001$; two-sample Wilcoxon $W$-test, $W = 62$) and \%TN ($p$-value $< 0.001$, $W = 102$) (Fig.
 \ref{fig:2}).
@@
 
-Differences in bulk densities between the two forest types resulted in a completely different distribution of C compared to the previous results based on \%OC alone (Table \ref{tab:1}).
+Differences in bulk densities between the two forest types, multiplied by \% organic carbon, resulted in a completely different distribution of C compared to the previous results based on \%OC alone (Table \ref{tab:1}).
 A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}).
-The differences in TOC between basin and fringe mangrove types are significant according to the one-way analysis of variance test ($p$-value $= 0.032$, $F = 4.88$, 48 d.f.).
+Summed across the profile, fringe mangroves had a TOC of 387.77 $\pm$ 125.77 Mg C ha$^{-1}$, and basing mangroves an average TOC of 665.41 $\pm$ 434.65 Mg C ha$^{-1}$.
+Given this large variability, we did not find enough evidence to reject the hypothesis that TOC from these two forest types are sampled from the same distribution  ($p$-value $= 0.47$, $W = 350$).
 TOC decreased with soil depth within the first four (basin) and three (fringe) layers.

### 2018-06-28 Reviews/Response2.tex
Addressed comments form reviewer 5

@@
 
+{\color{blue} Thanks for pointing this our.
+We originally applied an ANOVA test without checking the normality assumption.
+After checking for normality using the Shapiro-Wilk test, we found that the normally assumption was not met by our data, even after trying different transformations.
+Therefore, we decided to apply a non-parametric test, the Mann-Whitney U test to determine whether the two independent samples from the fringe and basin mangroves are from the same distributions.
+After running this test for all the measured variables, we found the same results as with the initial comparison using ANOVA.
+Therefore, our conclusions are unchanged despite applying a different statistical test.
+}
+
 {\it Why did you not add data on N/C ratio (or at least C/N ratio) as the reviewer 1 has suggested?

### 2018-06-28 Code_and_data/reprod.R

@@ -12,24 +12,47 @@ C3_80_85<-read.csv("C3_80-85.csv")
 
 #######################################################
 # Statistical comparisons
-mangroveOnly=rbind(Ergebnisse[Ergebnisse$Type=="Fringe",], Ergebnisse[Ergebnisse$Type=="Basin",]) #Exclude data from the end members
+mangroveOnly=rbind(Ergebnisse[Ergebnisse$Type=="Fringe",], Ergebnisse[Ergebnisse$Type=="Basin",]) # Exclude data from the end members
+mangAlFe=Al_Fe[Al_Fe[,"Type"]=="Fringe" | Al_Fe[,"Type"]=="Basin",c("Type","Al_Sum", "Fe_Sum")] # Exclude data from the end members
 
 # Comparison for Organic carbon
-anova(lm(Corg~Type, data=mangroveOnly))
+shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Fringe","Corg"]) #Normality test
+shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Basin","Corg"])
+
+wilcox.test(Corg~Type, data=mangroveOnly, exact=FALSE) #Non-parametric test for comparing whether samples are from the same distribution. For two samples, the test is know as 'Mann-Whitney test'.
 
 # Comparison for total nitrogen
-anova(lm(Ntotal~Type, data=mangroveOnly))
+shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Fringe","Ntotal"])
+shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Basin","Ntotal"])
+
+wilcox.test(Ntotal~Type, data=mangroveOnly, exact=FALSE)
 
 # Comparison TOC
-anova(lm(TOC~Type, data=TOC))
+# Data must be aggregated across depth for each plot
+plotTOC=tapply(TOC[,"TOC"], TOC[,"Plot_Number"], FUN=sum)
 
-# Comparison 13C
-anova(lm(d13C~Type, data=mangroveOnly))
+mean(plotTOC[c("P21", "R1", "R4", "R5", "R6")]) # Fringe plots
+sd(plotTOC[c("P21", "R1", "R4", "R5", "R6")])
+
[…]

### 2018-06-29 Manuscript/carbonStabilizationMangrovesV2.tex
Worked on manuscript text

@@
 
+Different techniques have been used in the past to determine different sources of organic carbon in water saturated ecosystems.
+For instance, the $\delta^{13}$C isotope has been used to explore different sources of carbon by source partitioning analyses \citep{Bouillion, Spohn2012, Spohn2013}.
+Also, measurements of the mineralogical composition of the sediments can also give some clues on whether the minerals present in an area are from terrestrial or marine origin.
+
 Here, our main objective was to determine the origin of the relatively high carbon levels in these soils using elemental and isotopic analyses of carbon as well as analyses of the soils' mineralogy in concert with Al- and Fe-oxide measurements.
-Analyses of stable isotopes are particularly useful to identify the origin of carbon in soils of coastal areas \citep{Bouillion, Spohn2012, Spohn2013}.
 In particular, we expect that: 1) the isotopic composition ($\updelta ^{13}$C) of soil carbon provides information on whether riverine sediments are a main source of C in Cisipat\'a bay, Colombia; and  2) the mineralogical composition of the sediments provides additional information on the origin of the soil carbon and the potential that it is mostly stabilized on the surface of Fe- and Al-oxides.
@@
 \ref{fig:1}).
-Generally, cores without layer changes were divided into sections every 20 cm to explore differences of carbon concentration with depth.
+These river and beach sediment samples were used as end-members in our isotopic analysis.
+Generally, cores without layer changes were divided into sections every 20 cm to explore differences in carbon concentration with depth.
 We divided conspicuous layer changes at the boundary, except for the sand and river sediment samples that were analyzed as one single sample for mineralogical and elemental composition.
@@
 We determined each mineral phase using the powder diffraction file (PDF) data.
-We also conducted a Rietveld refinement using the Topas software for XRD analysis (Bruker Corporation).
+We also conducted a Rietveld refinement analysis using the Topas software (Bruker Corporation).
 To define the type of clays included in the samples, we further measured each sample before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels.
@@
 \subsection{Data analysis}
-We performed statistical tests to determine differences in organic carbon concentrations, total nitrogen, $\delta^{13}$C, Fe- and Al-oxides, among the fringe and basin mangroves.
+We performed statistical tests to determine differences in organic carbon concentrations, total organic carbon, $\delta^{13}$C, and Fe- and Al-oxides, among the fringe and basin mangroves.
 After an initial test for normality (Shapiro-Wilk test), we found little evidence to support the normality assumption required in common tools such as ANOVA or t-tests.
-We therefore, performed the non-parametric Mann-Whitney U test, which is equivalent to the two-sample Wilcoxon test \citep{Hollander2015} under the null hypothesis that the samples from the fringe and basin mangroves  differ by a shift location $\mu =0$, and the alternative hypothesis that they different by some other shift location.
+We therefore, performed the non-parametric Mann-Whitney U test, which is equivalent to the two-sample Wilcoxon test \citep{Hollander2015} under the null hypothesis that the samples from the fringe and basin mangroves  differ by a shift location $\mu =0$, and the alternative hypothesis that they are different by some other shift location.
 In other words, the test helps to determine whether the samples from the two different mangrove types belong to the same statistical distribution as stated in the null hypothesis.
@@
 
-Differences in bulk densities between the two forest types, multiplied by \% organic carbon, resulted in a completely different distribution of C compared to the previous results based on \%OC alone (Table \ref{tab:1}).
+The different bulk densities between the two forest types, multiplied by their respective \% organic carbon, resulted in a completely different distribution of TOC compared to the previous results based on \%OC alone (Table \ref{tab:1}).
 A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}).
 Summed across the profile, fringe mangroves had a TOC of 387.77 $\pm$ 125.77 Mg C ha$^{-1}$, and basing mangroves an average TOC of 665.41 $\pm$ 434.65 Mg C ha$^{-1}$.
-Given this large variability, we did not find enough evidence to reject the hypothesis that TOC from these two forest types are sampled from the same distribution  ($p$-value $= 0.47$, $W = 350$).
+Given this large variability, and deespite average TOC being almost twice as high in the basing than in the fringe mangroves, we did not find enough evidence to reject the hypothesis that TOC from these two forest types are sampled from the same distribution  ($p$-value $= 0.47$, $W = 350$).
 TOC decreased with soil depth within the first four (basin) and three (fringe) layers.
@@
 We found no identifiable continuous trend between $^{13}$C values and depth.
-However, differences in $^{13}$C mean values between fringe and basin mangroves were statistically significant ($p$-value $< 0.001$, $F = 24.35$, 49 d.f.).
+However, differences in $^{13}$C mean values between fringe and basin mangroves suggest two different distributions ($p$-value $< 0.001$, $W = 542$).
 
-There were no differences between basin and fringe mangroves in terms of their composition of the mineral fraction (Table \ref{tab:2}).
+There were no differences between basin and fringe mangroves in terms of the composition of their mineral fraction (Table \ref{tab:2}).
 The proportion of mineral soil is larger in basin than in fringe mangroves.
@@
 \ref{fig:6})}.
-Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $< 0.001$, $F = 16.35$, 39 d.f.).
-Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $F = 33.45$, 39 d.f.).
+Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $= 0.002$, $W = 325$).
+Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $W = 364$).
 Basin soil samples had two times higher concentrations of Al (Al$_d$+Al$_o$) than fringe soils, and five times higher Fe values.
@@
 Our results confirmed previously observed differences in \%OC, bulk density and \%TOC between the basin and the fringe mangrove soils \citep{Bolivar2015}.
-Furthermore, our measurements of carbon isotopes and mineralogy helped us to establish the potential origin of the carbon stored in both mangrove types.
-In the following we will discuss these difference and the potential implications of our findings.
+Furthermore, our measurements of carbon isotopes and mineralogy helped us to explore the potential origin of the carbon stored in both mangrove types.
+In the following, we will discuss these difference and the potential implications of our findings.
 

### 2018-06-29 Reviews/Response2.tex
Worked on manuscript text

@@
 
-{\color{blue} Thanks for pointing this our.
+{\color{blue} Thanks for pointing this out.
 We originally applied an ANOVA test without checking the normality assumption.
@@
 
+{\color{blue} Organic C concentrations and total N show a very high correlation among each other (0.96 Pearson correlation coefficient), and the C:N ratio does not differ between the two forest types.
+Therefore, the information in the total N variable is redundant with respect to the information in organic C.
+Therefore, we think that adding an analysis of the C:N ratio does not contribute much to the story, as opposed to using other independent variables such as the isotopic composition of carbon, and the mineralogical composition of the sediments.
+}
+
 \vspace{2em}

### 2018-07-02 Reviews/Response2.tex
Prepared revision and resubmitted to journal

@@
 Thank you very much for your consideration of our manuscript and for the opportunity to submit a revised version.
-We made several changes to the manuscript based on the reviewers' comments.
-More prominently, we \dots
+We made several changes to the manuscript based on reviewers' comments.
+Most changes attempt at increasing clarity addressing reviewers' suggestions.
+In addition, we changed the statistical test used for comparisons, using now a non-parametric test that does not rely on the assumption of normality of the data.
+For details, see the answers to the comments below.
 
@@
 
+{\color{blue} We thank the reviewer for recognizing the value of this study and pointing out different issues such as lack of detailed descriptions and confidence in wording.
+We have made an effort to address all these issues (see answers below), and as a result the manuscript is now much improved.
+In some cases however, we think some of the issues raised by the reviewer are due to misunderstandings of the text.
+In these cases, we made an effort to present a more clear text.
+}
+
 \begin{enumerate}
@@
 
-{\color{blue} Habitat and environmental characteristics of the basing and fringe mangroves are described in the Mathrials and Methods section.
+{\color{blue} Habitat and environmental characteristics of the basing and fringe mangroves are described in the Materials and Methods section.
 Here in the introduction, we are describing the main motivation of the study and therefore it is not the most appropriate place to describe the study site.
@@
 What is the rationale supporting this line of inquiry - and most importantly - where are the literature sources to validate it?}
+
+{\color{blue} We added a new paragraph giving this rationale. }
 
@@
 
-{\color{blue} The rationale is presented in the introduction, where we mention that we use analyses of the structure and composition of the clay minerals to help in determining the origin of the stored carbon.}
+{\color{blue} The rationale is presented now in the introduction, where we mention that we use analyses of the structure and composition of the clay minerals to help in determining the origin of the stored carbon.}
 
@@
 Therefore the Tukey post-hoc test is not needed.
+However, given in the new version of the manuscript we do not use ANOVA anymore for comparisons of means.
+Since the data was not normally distributed, we use now a Wilcoxon test to compare the distribution of the data.
+Description of statistical analyses was changed accordingly in the text.
 }
@@
 We do not think we need to give a complex argumentation for this result since it is almost trivial.
-However, in the new version we point the reader to Table 1 to see this in more detail.
+However, in the new version we explain this result better and point the reader to Table 1 to see this in more detail.
 }
@@
 
+{\color{blue} We reworded this paragraph.
+The point is simply that the results helped us to make an interpretation of the origin of the carbon, but this is analyzed only in the discussion, independent of the results section where se state the result without interpret them.
+}
+
 \item {\it Ln 236 Reference to support this claim? Surely you were sampling root-free soils? Or did your soil samples also contain live root biomass?}
@@
 {\color{blue} Yes, this happens in most soils, but in mangroves bulk densities increase strongly due to hydrostatic pressure.
-However, we agree in that this sentence is not very relevant for the discussion on differences in bulk density among the two mangrove types.}
+However, we agree in that this sentence is not very relevant for the discussion on differences in bulk density among the two mangrove types, and therefore we removed it.}
 
@@
 Without such validating information, until now, what you have done has remained pretty much meaningless to the reader.
+
+{\color{blue} Some of this information was added to the introduction.}
 
@@
 or ...' , 'which would underpin', etc.
-These are all hedged statements that show our interpretation of the results and our favoring for a specific interpretation, but away from a definite true.}
+These are all hedged statements that show our interpretation of the results and our favoring for a specific interpretation, but away from a definite true.
+\\ Despite our disagreement on this matter, we did try to change the language in several parts of the manuscript to make our statements less `definite'.}
 
@@
 
-{\color{blue} Organic C concentrations and total N show a very high correlation among each other (0.96 Pearson correlation coefficient), and the C:N ratio does not differ between the two forest types.
+{\color{blue} Organic C concentrations and total N show a very high correlation among each other (0.96 correlation coefficient), and the C:N ratio does not differ between the two forest types.
 Therefore, the information in the total N variable is redundant with respect to the information in organic C.

### 2018-07-02 Code_and_data/reprod.R

@@ -27,18 +27,6 @@ shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Basin","Ntotal"])
 
 wilcox.test(Ntotal~Type, data=mangroveOnly, exact=FALSE)
 
-# Comparison TOC
-# Data must be aggregated across depth for each plot
-plotTOC=tapply(TOC[,"TOC"], TOC[,"Plot_Number"], FUN=sum)
-
-mean(plotTOC[c("P21", "R1", "R4", "R5", "R6")]) # Fringe plots
-sd(plotTOC[c("P21", "R1", "R4", "R5", "R6")])
-
-mean(plotTOC[c("P16", "C1", "C2", "C3", "C4")]) # Basin plots
-sd(plotTOC[c("P16", "C1", "C2", "C3", "C4")])
-
-wilcox.test(TOC~Type, data=TOC, exact=FALSE)
-
  # Comparison 13C
 shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Fringe","d13C"])
 shapiro.test(mangroveOnly[mangroveOnly[,"Type"]=="Basin","d13C"])
@@ -223,15 +211,13 @@ dev.off()
 
 # Mean values with propagated uncertainty
 # For TOC
-B=c(295.61,161.99,68.35,65.21,69.78)
-sum(B)
-sdB=c(203.77,147.07,40.95,30.58,37.60)
-sqrt(sum(sdB^2)/length(sdB))
+meanTOCdepth=tapply(TOC[,"TOC"], list(TOC[,"Type"], TOC[,"Depth_adjusted"]),FUN=mean)
+sdTOCdepth=tapply(TOC[,"TOC"], list(TOC[,"Type"], TOC[,"Depth_adjusted"]),FUN=sd)
 
-Fr=c(100.24,62.79,51.57,105.44)
-sdF=c(9.03,22.56,51.57,105.44)
-sum(Fr)
-sqrt(sum(sdF^2)/length(sdF))
+sumTOC=rowSums(meanTOCdepth, na.rm=TRUE)
+round(sumTOC, 2)
+sdTOC=apply(sdTOCdepth, 1, FUN=function(x){sqrt(sum(x^2,na.rm=TRUE))})
+round(sdTOC,2)
 
 # For Bulk density
 round(mean(TOC[TOC[,"Type"]=="Basin","BulkDens"]), 2)
